---
title: "Zika Pipeline Status Report"
subtitle: "Stages 1-3 MVP diagnostics and current outputs"
author: "INFLUX Impact"
date: last-modified
format:
  html:
    toc: true
    number-sections: true
    code-fold: show
execute:
  echo: false
  warning: false
  message: false
---

## Scope

This report summarizes the current state of the Zika news-mining MVP pipeline in `zika/`:

1. Google News retrieval and cached manifest status
2. Headline ensemble filtering diagnostics and validation assets
3. Full-text retrieval status

The report is designed to read generated artifacts directly from disk and surface incomplete or invalid states explicitly.

::: {.callout-warning}
Stage 2 outputs should only be treated as complete when both model outputs are present and both `openai_error` and `local_error` are empty.
:::

```{r}
#| label: setup-r
pacman::p_load(
  rio, # File import
  here, # File locator
  arrow,
  reticulate,
  httr,
  jsonlite,
  stringr,
  tidyverse # data management + ggplot2 graphics
)

use_virtualenv(here::here(".venv"))
# arrow::install_pyarrow() # need to install pip
Sys.setenv(RETICULATE_PYTHON = here::here(".venv"))

repl_python()

find_repo_root <- function(start = getwd()) {
  current <- normalizePath(start, winslash = "/", mustWork = TRUE)
  while (TRUE) {
    if (
      dir.exists(file.path(current, "zika")) ||
        dir.exists(file.path(current, "src"))
    ) {
      return(current)
    }
    parent <- dirname(current)
    if (identical(parent, current)) {
      stop("Could not locate repository root.")
    }
    current <- parent
  }
}

ROOT <- find_repo_root()
ZIKA <- file.path(ROOT, "zika")

safe_read_csv <- function(path) {
  if (!file.exists(path)) {
    return(tibble())
  }
  suppressMessages(readr::read_csv(path, show_col_types = FALSE))
}

artifact_info <- function(path_label, path) {
  exists <- file.exists(path)
  rel_path <- if (startsWith(path, paste0(ROOT, "/"))) {
    substring(path, nchar(ROOT) + 2L)
  } else {
    path
  }
  tibble(
    artifact = path_label,
    path = rel_path,
    exists = exists,
    size_kb = if (exists) round(file.info(path)$size / 1024, 1) else NA_real_,
    modified_at = if (exists) {
      as.character(file.info(path)$mtime)
    } else {
      NA_character_
    }
  )
}

`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0) y else x
}

manifest_path <- file.path(ZIKA, "data", "raw", "rss", "zika_rss_manifest.csv")
rss_path <- file.path(ZIKA, "data", "intermediate", "zika_rss_raw.csv")
scored_path <- file.path(
  ZIKA,
  "data",
  "intermediate",
  "zika_headlines_scored.csv"
)
validation_path <- file.path(
  ZIKA,
  "data",
  "validation",
  "zika_headline_validation_sample.csv"
)
metrics_json_path <- file.path(
  ZIKA,
  "data",
  "validation",
  "zika_headline_metrics.json"
)
fulltext_path <- file.path(ZIKA, "data", "final", "zika_fulltext.csv")
fulltext_failed_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_fulltext_failed.csv"
)

manifest_df <- safe_read_csv(manifest_path)
rss_df <- safe_read_csv(rss_path)
scored_df <- safe_read_csv(scored_path)
validation_df <- safe_read_csv(validation_path)
fulltext_df <- safe_read_csv(fulltext_path)
fulltext_failed_df <- safe_read_csv(fulltext_failed_path)
metrics_payload <- if (file.exists(metrics_json_path)) {
  jsonlite::fromJSON(metrics_json_path, simplifyVector = TRUE)
} else {
  list()
}

completed_score_mask <- function(df) {
  if (!nrow(df)) {
    return(logical())
  }
  openai_ok <- !is.na(df$openai_label) &
    df$openai_label %in% c("relevant", "irrelevant", "unsure")
  local_ok <- !is.na(df$local_label) &
    df$local_label %in% c("relevant", "irrelevant", "unsure")
  action_ok <- !is.na(df$final_action) &
    df$final_action %in% c("keep", "drop", "review")
  openai_err <- if ("openai_error" %in% names(df)) {
    trimws(coalesce(df$openai_error, ""))
  } else {
    ""
  }
  local_err <- if ("local_error" %in% names(df)) {
    trimws(coalesce(df$local_error, ""))
  } else {
    ""
  }
  openai_ok & local_ok & action_ok & openai_err == "" & local_err == ""
}

completed_scored_df <- if (nrow(scored_df)) {
  scored_df[completed_score_mask(scored_df), , drop = FALSE]
} else {
  tibble()
}
```

## Artifact Inventory

```{r}
bind_rows(
  artifact_info("Stage 1 manifest", manifest_path),
  artifact_info("Stage 1 normalized RSS", rss_path),
  artifact_info("Stage 2 scored headlines", scored_path),
  artifact_info("Stage 2 validation sample", validation_path),
  artifact_info("Stage 2 metrics JSON", metrics_json_path),
  artifact_info("Stage 3 full text", fulltext_path),
  artifact_info("Stage 3 failed full text", fulltext_failed_path)
)
```

## Stage 1 Retrieval

```{r}
if (nrow(manifest_df)) {
  manifest_df %>%
    mutate(status = coalesce(status, "missing")) %>%
    count(status, sort = TRUE, name = "windows")
} else {
  tibble(note = "No manifest found.")
}
```

```{r}
if (nrow(manifest_df)) {
  tibble(
    metric = c(
      "manifest_rows",
      "successful_windows",
      "empty_windows",
      "throttled_windows",
      "cached_windows"
    ),
    value = c(
      nrow(manifest_df),
      sum(manifest_df$status == "success", na.rm = TRUE),
      sum(manifest_df$status == "empty", na.rm = TRUE),
      sum(manifest_df$status == "throttled", na.rm = TRUE),
      sum(
        !is.na(manifest_df$cached_path) & trimws(manifest_df$cached_path) != "",
        na.rm = TRUE
      )
    )
  )
} else {
  tibble(note = "No manifest metrics available.")
}
```

```{r}
#| label: fig-stage1-status
#| fig-cap: "Stage 1 manifest status by window."
if (nrow(manifest_df)) {
  manifest_df %>%
    mutate(status = coalesce(status, "missing")) %>%
    count(status, sort = TRUE, name = "windows") %>%
    ggplot(aes(x = reorder(status, windows), y = windows, fill = status)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    labs(x = NULL, y = "Windows") +
    theme_minimal(base_size = 12)
}
```

```{r}
if (nrow(rss_df)) {
  tibble(
    metric = c(
      "normalized_rows",
      "unique_redirect_urls",
      "unique_guid",
      "date_min",
      "date_max"
    ),
    value = c(
      nrow(rss_df),
      dplyr::n_distinct(rss_df$google_news_redirect_url, na.rm = TRUE),
      dplyr::n_distinct(rss_df$guid, na.rm = TRUE),
      as.character(min(rss_df$date_window_start, na.rm = TRUE)),
      as.character(max(rss_df$date_window_end, na.rm = TRUE))
    )
  )
} else {
  tibble(note = "No normalized RSS table found.")
}
```

## Stage 2 Ensemble Filtering

```{r}
if (nrow(scored_df)) {
  tibble(
    metric = c(
      "headline_rows",
      "completed_rows",
      "incomplete_rows",
      "openai_error_rows",
      "local_error_rows"
    ),
    value = c(
      nrow(scored_df),
      nrow(completed_scored_df),
      nrow(scored_df) - nrow(completed_scored_df),
      sum(trimws(coalesce(scored_df$openai_error, "")) != "", na.rm = TRUE),
      sum(trimws(coalesce(scored_df$local_error, "")) != "", na.rm = TRUE)
    )
  )
} else {
  tibble(note = "No Stage 2 scored output found.")
}
```

```{r}
if (nrow(completed_scored_df)) {
  completed_scored_df %>%
    count(final_action, sort = TRUE, name = "rows")
} else {
  tibble(note = "No completed ensemble rows are currently available.")
}
```

```{r}
if (nrow(completed_scored_df)) {
  bind_rows(
    completed_scored_df %>%
      count(model = "OpenAI", label = openai_label, name = "rows"),
    completed_scored_df %>%
      count(model = "Local", label = local_label, name = "rows")
  )
} else {
  tibble(
    note = "Model label distributions are unavailable until Stage 2 completes cleanly."
  )
}
```

```{python}
from __future__ import annotations

from pathlib import Path
import pandas as pd


def find_repo_root(start: Path) -> Path:
    for path in [start, *start.parents]:
        if (path / "zika").exists():
            return path
    raise RuntimeError("Could not locate repository root.")


ROOT = find_repo_root(Path.cwd())
scored_path = ROOT / "zika" / "data" / "intermediate" / "zika_headlines_scored.csv"

if scored_path.exists():
    scored = pd.read_csv(scored_path, dtype=str).fillna("")
    error_frames = []
    for field in ["openai_error", "local_error"]:
        subset = (
            scored.loc[scored[field].str.strip() != "", field].value_counts().head(10)
        )
        if not subset.empty:
            error_frames.append(
                pd.DataFrame(
                    {
                        "field": field,
                        "error": subset.index,
                        "count": subset.values,
                    }
                )
            )

    if error_frames:
        pd.concat(error_frames, ignore_index=True)
    else:
        pd.DataFrame({"note": ["No provider errors recorded."]})
else:
    pd.DataFrame({"note": ["Scored CSV not found."]})
```

## Validation Assets

```{r}
if (nrow(validation_df)) {
  tibble(
    metric = c("validation_rows", "labeled_rows", "unlabeled_rows"),
    value = c(
      nrow(validation_df),
      sum(
        !is.na(validation_df$gold_label) &
          trimws(validation_df$gold_label) != "",
        na.rm = TRUE
      ),
      sum(
        is.na(validation_df$gold_label) |
          trimws(validation_df$gold_label) == "",
        na.rm = TRUE
      )
    )
  )
} else {
  tibble(note = "Validation sample is empty or absent.")
}
```

```{r}
if (length(metrics_payload)) {
  tibble(
    metric = c(
      "completed_rows",
      "incomplete_rows",
      "local_error_rows",
      "openai_error_rows"
    ),
    value = c(
      metrics_payload$completed_rows %||% NA,
      metrics_payload$incomplete_rows %||% NA,
      metrics_payload$errors$local_error_rows %||% NA,
      metrics_payload$errors$openai_error_rows %||% NA
    )
  )
} else {
  tibble(note = "Metrics JSON not found.")
}
```

## Stage 3 Full Text

```{r}
if (nrow(fulltext_df) || nrow(fulltext_failed_df)) {
  tibble(
    metric = c("successful_fulltext_rows", "failed_fulltext_rows"),
    value = c(nrow(fulltext_df), nrow(fulltext_failed_df))
  )
} else {
  tibble(note = "No full-text outputs are present yet.")
}
```

## Current Interpretation

```{r}
if (nrow(completed_scored_df) == 0 && nrow(scored_df) > 0) {
  tibble(
    finding = "Stage 2 is not currently complete.",
    detail = "The scored CSV exists, but no rows satisfy the completed ensemble criteria. Provider errors must be resolved before using validation outputs."
  )
} else if (nrow(completed_scored_df) > 0) {
  tibble(
    finding = "Stage 2 has completed rows available.",
    detail = "Validation and downstream full-text retrieval can proceed on retained URLs."
  )
} else {
  tibble(
    finding = "Pipeline artifacts are not yet available.",
    detail = "Run the pipeline stages first."
  )
}
```
