---
title: "Zika Pipeline Status Report"
subtitle: "Ensemble filtering and full-text retrieval results"
author: "INFLUX Impact"
date: last-modified
format:
  html:
    toc: true
    number-sections: false
    code-fold: show
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

## About this report

This report summarizes the result of experiments on applying ensemble learning to classifiy Zika news articles pipeline outputs. 
The report is organized into three main sections corresponding to the three stages of the pipeline:

1. Google News retrieval and cached manifest status
2. Headline ensemble filtering diagnostics using cloud OpenAI (gpt-5-nano) and local DeepSeek (deepseek-r1:14b) models
3. Full-text retrieval status

```{r}
#| label: setup
pacman::p_load(
  rio, # File import
  here, # File locator
  arrow,
  reticulate,
  httr,
  jsonlite,
  stringr,
  kableExtra,
  tidyverse # data management + ggplot2 graphics
)

use_virtualenv(here::here(".venv"))
# arrow::install_pyarrow() # need to install pip
Sys.setenv(RETICULATE_PYTHON = here::here(".venv"))

# repl_python()
```

```{r}
#| label: utils
find_repo_root <- function(start = getwd()) {
  current <- normalizePath(start, winslash = "/", mustWork = TRUE)
  while (TRUE) {
    if (
      dir.exists(file.path(current, "zika")) ||
        dir.exists(file.path(current, "src"))
    ) {
      return(current)
    }
    parent <- dirname(current)
    if (identical(parent, current)) {
      stop("Could not locate repository root.")
    }
    current <- parent
  }
}

ROOT <- find_repo_root()
ZIKA <- file.path(ROOT, "zika")

safe_read_csv <- function(path) {
  if (!file.exists(path)) {
    return(tibble())
  }
  suppressMessages(readr::read_csv(path, show_col_types = FALSE))
}

sanitize_json_text <- function(text) {
  text %>%
    str_replace_all("(?<![A-Za-z0-9_])NaN(?![A-Za-z0-9_])", "null") %>%
    str_replace_all("(?<![A-Za-z0-9_])-Infinity(?![A-Za-z0-9_])", "null") %>%
    str_replace_all("(?<![A-Za-z0-9_])Infinity(?![A-Za-z0-9_])", "null")
}

safe_read_json <- function(path) {
  if (!file.exists(path)) {
    return(list())
  }
  json_text <- readr::read_file(path) %>%
    sanitize_json_text()
  jsonlite::fromJSON(json_text, simplifyVector = TRUE)
}

artifact_info <- function(path_label, path) {
  exists <- file.exists(path)
  rel_path <- if (startsWith(path, paste0(ROOT, "/"))) {
    substring(path, nchar(ROOT) + 2L)
  } else {
    path
  }
  tibble(
    artifact = path_label,
    path = rel_path,
    exists = exists,
    size_kb = if (exists) round(file.info(path)$size / 1024, 1) else NA_real_,
    modified_at = if (exists) {
      as.character(file.info(path)$mtime)
    } else {
      NA_character_
    }
  )
}

`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0) y else x
}
```

```{r}
#| label: load-artifacts
manifest_path <- file.path(ZIKA, "data", "raw", "rss", "zika_rss_manifest.csv")
rss_path <- file.path(ZIKA, "data", "intermediate", "zika_rss_raw.csv")
scored_path <- file.path(
  ZIKA,
  "data",
  "intermediate",
  "zika_headlines_scored.csv"
)
validation_path <- file.path(
  ZIKA,
  "data",
  "validation",
  "zika_headline_validation_sample.csv"
)
metrics_json_path <- file.path(
  ZIKA,
  "data",
  "validation",
  "zika_headline_metrics.json"
)
fulltext_path <- file.path(ZIKA, "data", "final", "zika_fulltext.csv")
fulltext_failed_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_fulltext_failed.csv"
)
bertopic_doc_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_document_topics.csv"
)
bertopic_topic_info_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_topic_info.csv"
)
bertopic_summary_json_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_summary.json"
)
bertopic_compare_csv_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_model_comparison.csv"
)
bertopic_compare_json_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_model_comparison.json"
)
bertopic_modular_csv_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_modular_analysis.csv"
)
bertopic_modular_json_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_modular_analysis.json"
)
bertopic_modular_topics_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_modular_topics.csv"
)
bertopic_stage_timing_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_stage_timings.csv"
)
bertopic_topic_match_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_topic_matches.csv"
)
bertopic_minilm_doc_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_all_minilm_l6_v2_document_topics.csv"
)
bertopic_minilm_topic_info_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_all_minilm_l6_v2_topic_info.csv"
)
bertopic_bge_doc_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_baai_bge_base_en_v1_5_document_topics.csv"
)
bertopic_bge_topic_info_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_baai_bge_base_en_v1_5_topic_info.csv"
)
bertopic_impact_doc_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_impact_document_topics.csv"
)
bertopic_impact_topic_info_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_impact_topic_info.csv"
)
bertopic_impact_summary_json_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_impact_summary.json"
)
bertopic_impact_projection_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_impact_projection.csv"
)
bertopic_impact_stage_timing_path <- file.path(
  ZIKA,
  "data",
  "final",
  "zika_bertopic_impact_stage_timings.csv"
)

manifest_df <- safe_read_csv(manifest_path)
rss_df <- safe_read_csv(rss_path)
scored_df <- safe_read_csv(scored_path)
validation_df <- safe_read_csv(validation_path)
fulltext_df <- safe_read_csv(fulltext_path)
fulltext_failed_df <- safe_read_csv(fulltext_failed_path)
bertopic_doc_df <- safe_read_csv(bertopic_doc_path)
bertopic_topic_info_df <- safe_read_csv(bertopic_topic_info_path)
bertopic_compare_df <- safe_read_csv(bertopic_compare_csv_path)
bertopic_modular_df <- safe_read_csv(bertopic_modular_csv_path)
bertopic_modular_topics_df <- safe_read_csv(bertopic_modular_topics_path)
bertopic_stage_timing_df <- safe_read_csv(bertopic_stage_timing_path)
bertopic_topic_match_df <- safe_read_csv(bertopic_topic_match_path)
bertopic_minilm_doc_df <- safe_read_csv(bertopic_minilm_doc_path)
bertopic_minilm_topic_info_df <- safe_read_csv(bertopic_minilm_topic_info_path)
bertopic_bge_doc_df <- safe_read_csv(bertopic_bge_doc_path)
bertopic_bge_topic_info_df <- safe_read_csv(bertopic_bge_topic_info_path)
bertopic_impact_doc_df <- safe_read_csv(bertopic_impact_doc_path)
bertopic_impact_topic_info_df <- safe_read_csv(bertopic_impact_topic_info_path)
bertopic_impact_projection_df <- safe_read_csv(bertopic_impact_projection_path)
bertopic_impact_stage_timing_df <- safe_read_csv(
  bertopic_impact_stage_timing_path
)
metrics_payload <- if (file.exists(metrics_json_path)) {
  safe_read_json(metrics_json_path)
} else {
  list()
}
bertopic_summary_payload <- safe_read_json(bertopic_summary_json_path)
bertopic_compare_payload <- safe_read_json(bertopic_compare_json_path)
bertopic_modular_payload <- safe_read_json(bertopic_modular_json_path)
bertopic_impact_payload <- safe_read_json(bertopic_impact_summary_json_path)

completed_score_mask <- function(df) {
  if (!nrow(df)) {
    return(logical())
  }
  openai_ok <- !is.na(df$openai_label) &
    df$openai_label %in% c("relevant", "irrelevant", "unsure")
  local_ok <- !is.na(df$local_label) &
    df$local_label %in% c("relevant", "irrelevant", "unsure")
  action_ok <- !is.na(df$final_action) &
    df$final_action %in% c("keep", "drop", "review")
  openai_err <- if ("openai_error" %in% names(df)) {
    trimws(coalesce(df$openai_error, ""))
  } else {
    ""
  }
  local_err <- if ("local_error" %in% names(df)) {
    trimws(coalesce(df$local_error, ""))
  } else {
    ""
  }
  openai_ok & local_ok & action_ok & openai_err == "" & local_err == ""
}

completed_scored_df <- if (nrow(scored_df)) {
  scored_df[completed_score_mask(scored_df), , drop = FALSE]
} else {
  tibble()
}
```

## Output files directory and status

```{r}
bind_rows(
  artifact_info("Stage 1 manifest", manifest_path),
  artifact_info("Stage 1 normalized RSS", rss_path),
  artifact_info("Stage 2 scored headlines", scored_path),
  artifact_info("Stage 2 validation sample", validation_path),
  artifact_info("Stage 2 metrics JSON", metrics_json_path),
  artifact_info("Stage 3 full text", fulltext_path),
  artifact_info("Stage 3 failed full text", fulltext_failed_path),
  artifact_info("Stage 4 BERTopic document topics", bertopic_doc_path),
  artifact_info("Stage 4 BERTopic topic info", bertopic_topic_info_path),
  artifact_info("Stage 4 BERTopic summary JSON", bertopic_summary_json_path),
  artifact_info(
    "Stage 4 BERTopic impact document topics",
    bertopic_impact_doc_path
  ),
  artifact_info(
    "Stage 4 BERTopic impact topic info",
    bertopic_impact_topic_info_path
  ),
  artifact_info(
    "Stage 4 BERTopic impact summary JSON",
    bertopic_impact_summary_json_path
  ),
  artifact_info(
    "Stage 4 BERTopic impact projection",
    bertopic_impact_projection_path
  ),
  artifact_info(
    "Stage 4 BERTopic impact stage timings",
    bertopic_impact_stage_timing_path
  ),
  artifact_info("Stage 4 BERTopic comparison CSV", bertopic_compare_csv_path),
  artifact_info("Stage 4 BERTopic comparison JSON", bertopic_compare_json_path),
  artifact_info(
    "Stage 4 BERTopic modular analysis CSV",
    bertopic_modular_csv_path
  ),
  artifact_info(
    "Stage 4 BERTopic modular analysis JSON",
    bertopic_modular_json_path
  ),
  artifact_info(
    "Stage 4 BERTopic modular topics CSV",
    bertopic_modular_topics_path
  ),
  artifact_info(
    "Stage 4 BERTopic stage timings CSV",
    bertopic_stage_timing_path
  ),
  artifact_info("Stage 4 BERTopic topic matches", bertopic_topic_match_path)
) %>%
  select(artifact, path, exists) %>%
  kableExtra::kable()
```

## Step 1 Retrieval

::: {.callout-note}
The manifest currently spans `r if (nrow(manifest_df)) min(manifest_df$window_start, na.rm = TRUE) else NA_character_` to `r if (nrow(manifest_df)) max(manifest_df$window_end, na.rm = TRUE) else NA_character_`.
:::

Stage 1 is implemented with a two-step workflow:

1. **Network retrieval** builds daily Google News search windows, checks the manifest, queries only unfinished windows, and writes one cached JSON payload per window.
2. **Parsing and normalization** re-reads cached payloads with `status %in% c("success", "empty")`, flattens the feed entries into a table, and applies deterministic deduplication before writing the final CSV and Parquet outputs.

The retrieval script is therefore optimized for interruption and restart, while the parsing script is optimized for fast reruns from cache.

### How the retrieval step works

1. Build daily windows from the configured query period.
2. Use the manifest as the authoritative checkpoint to decide whether a window is already complete, due for retry, or still pending.
3. Query Google News using the English / US edition configuration and persist the raw JSON payload to `zika/data/raw/rss/`.
4. Detect throttling from HTTP status and Google “Sorry” pages, then stop early instead of continuing to hammer adjacent windows.
5. Update the manifest after each attempt with timestamps, attempt counts, HTTP status, cache path, and error summary.
6. Parse only cached `success` and `empty` windows into a flat table.
7. Apply deterministic deduplication before writing the final Stage 1 CSV and Parquet outputs.

```{r}
#| label: stage1-helpers
retrieval_log_path <- file.path(ZIKA, "logs", "01_retrieve_google_rss.log")
parse_log_path <- file.path(ZIKA, "logs", "01_parse_google_rss.log")

parse_stage1_log <- function(path) {
  if (!file.exists(path)) {
    return(tibble())
  }

  readLines(path, warn = FALSE) %>%
    tibble(raw = .) %>%
    mutate(line_id = row_number()) %>%
    filter(str_detect(raw, "^\\d{4}-\\d{2}-\\d{2}")) %>%
    mutate(
      timestamp_txt = str_extract(
        raw,
        "^\\d{4}-\\d{2}-\\d{2} \\d{2}:\\d{2}:\\d{2}"
      ),
      timestamp = as.POSIXct(
        timestamp_txt,
        format = "%Y-%m-%d %H:%M:%S",
        tz = "Europe/Stockholm"
      ),
      message = str_trim(
        str_replace(
          raw,
          "^\\d{4}-\\d{2}-\\d{2}[^|]*\\|\\s*[A-Z]+\\s*\\|\\s*",
          ""
        )
      )
    )
}

pair_stage1_sessions <- function(
  starts_df,
  summaries_df,
  type = c("retrieval", "parse")
) {
  type <- match.arg(type)

  if (!nrow(starts_df)) {
    return(tibble())
  }

  sessions <- vector("list", nrow(starts_df))

  for (i in seq_len(nrow(starts_df))) {
    next_start_line <- if (i < nrow(starts_df)) {
      starts_df$line_id[i + 1]
    } else {
      Inf
    }
    hit <- summaries_df %>%
      filter(line_id > starts_df$line_id[i], line_id < next_start_line) %>%
      slice_head(n = 1)

    if (type == "retrieval") {
      sessions[[i]] <- tibble(
        run_started = starts_df$run_started[i],
        configured_windows = starts_df$configured_windows[i],
        configured_start = starts_df$configured_start[i],
        configured_end = starts_df$configured_end[i],
        backend = starts_df$backend[i],
        run_finished = if (nrow(hit)) hit$run_finished[1] else as.POSIXct(NA),
        duration_seconds = if (nrow(hit)) {
          as.numeric(difftime(
            hit$run_finished[1],
            starts_df$run_started[i],
            units = "secs"
          ))
        } else {
          NA_real_
        },
        status = if (nrow(hit)) "completed" else "interrupted",
        completed_windows = if (nrow(hit)) {
          hit$completed_windows[1]
        } else {
          NA_real_
        },
        success_fetches = if (nrow(hit)) hit$success_fetches[1] else NA_real_,
        empty_fetches = if (nrow(hit)) hit$empty_fetches[1] else NA_real_,
        failed_fetches = if (nrow(hit)) hit$failed_fetches[1] else NA_real_
      )
    } else {
      sessions[[i]] <- tibble(
        run_started = starts_df$run_started[i],
        cached_windows = starts_df$cached_windows[i],
        manifest_path = starts_df$manifest_path[i],
        run_finished = if (nrow(hit)) hit$run_finished[1] else as.POSIXct(NA),
        duration_seconds = if (nrow(hit)) {
          as.numeric(difftime(
            hit$run_finished[1],
            starts_df$run_started[i],
            units = "secs"
          ))
        } else {
          NA_real_
        },
        status = if (nrow(hit)) "completed" else "interrupted",
        parsed_rows = if (nrow(hit)) hit$parsed_rows[1] else NA_real_,
        deduped_rows = if (nrow(hit)) hit$deduped_rows[1] else NA_real_,
        missing_cache = if (nrow(hit)) hit$missing_cache[1] else NA_real_
      )
    }
  }

  bind_rows(sessions)
}

format_elapsed <- function(seconds) {
  if (is.na(seconds) || !is.finite(seconds)) {
    return(NA_character_)
  }
  seconds <- round(seconds)
  hours <- seconds %/% 3600
  minutes <- (seconds %% 3600) %/% 60
  secs <- seconds %% 60

  if (hours > 0) {
    sprintf("%dh %02dm %02ds", hours, minutes, secs)
  } else if (minutes > 0) {
    sprintf("%dm %02ds", minutes, secs)
  } else {
    sprintf("%ds", secs)
  }
}

retrieval_log_df <- parse_stage1_log(retrieval_log_path)
parse_log_df <- parse_stage1_log(parse_log_path)

retrieval_starts <- retrieval_log_df %>%
  filter(str_detect(message, "^Stage 1 retrieval configured")) %>%
  transmute(
    line_id,
    run_started = timestamp,
    configured_windows = as.integer(str_match(
      message,
      "configured for (\\d+) windows"
    )[, 2]),
    configured_start = str_match(
      message,
      "windows from (\\d{4}-\\d{2}-\\d{2}) to"
    )[, 2],
    configured_end = str_match(message, "to (\\d{4}-\\d{2}-\\d{2}) using")[, 2],
    backend = str_match(message, "backend=([A-Za-z0-9_:-]+)$")[, 2]
  )

retrieval_summaries <- retrieval_log_df %>%
  filter(str_detect(message, "^Retrieval summary")) %>%
  transmute(
    line_id,
    run_finished = timestamp,
    completed_windows = as.integer(str_match(message, "completed=(\\d+)")[, 2]),
    success_fetches = as.integer(str_match(message, "success=(\\d+)")[, 2]),
    empty_fetches = as.integer(str_match(message, "empty=(\\d+)")[, 2]),
    failed_fetches = as.integer(str_match(message, "failed=(\\d+)")[, 2])
  )

parse_starts <- parse_log_df %>%
  filter(str_detect(message, "^Parsing \\d+ cached windows")) %>%
  transmute(
    line_id,
    run_started = timestamp,
    cached_windows = as.integer(str_match(
      message,
      "^Parsing (\\d+) cached windows"
    )[, 2]),
    manifest_path = str_match(message, "manifest (.+)$")[, 2]
  )

parse_summaries <- parse_log_df %>%
  filter(str_detect(message, "^Parsing summary")) %>%
  transmute(
    line_id,
    run_finished = timestamp,
    parsed_rows = as.integer(str_match(message, "parsed_rows=(\\d+)")[, 2]),
    deduped_rows = as.integer(str_match(message, "deduped_rows=(\\d+)")[, 2]),
    missing_cache = as.integer(str_match(message, "missing_cache=(\\d+)")[, 2])
  )

retrieval_sessions <- pair_stage1_sessions(
  retrieval_starts,
  retrieval_summaries,
  type = "retrieval"
) %>%
  mutate(
    cached_reused = if_else(
      !is.na(completed_windows),
      completed_windows - success_fetches - empty_fetches,
      NA_real_
    ),
    elapsed = vapply(duration_seconds, format_elapsed, character(1))
  )

parse_sessions <- pair_stage1_sessions(
  parse_starts,
  parse_summaries,
  type = "parse"
) %>%
  mutate(elapsed = vapply(duration_seconds, format_elapsed, character(1)))

latest_retrieval <- retrieval_sessions %>%
  filter(status == "completed") %>%
  arrange(desc(run_started)) %>%
  slice_head(n = 1)

latest_parse <- parse_sessions %>%
  filter(status == "completed") %>%
  arrange(desc(run_started)) %>%
  slice_head(n = 1)

stage1_overview <- if (nrow(manifest_df)) {
  tibble(
    indicator = c(
      "manifest_rows",
      "manifest_unique_windows",
      "manifest_window_start",
      "manifest_window_end",
      "successful_windows",
      "empty_windows",
      "failed_or_throttled_windows",
      "cached_payloads_present",
      "final_rows_after_deduplication",
      "unique_redirect_urls",
      "unique_guid"
    ),
    value = c(
      nrow(manifest_df),
      n_distinct(paste(manifest_df$window_start, manifest_df$window_end)),
      min(manifest_df$window_start, na.rm = TRUE),
      max(manifest_df$window_end, na.rm = TRUE),
      sum(manifest_df$status == "success", na.rm = TRUE),
      sum(manifest_df$status == "empty", na.rm = TRUE),
      sum(manifest_df$status %in% c("failed", "throttled"), na.rm = TRUE),
      sum(trimws(coalesce(manifest_df$cached_path, "")) != "", na.rm = TRUE),
      nrow(rss_df),
      if (nrow(rss_df)) {
        n_distinct(rss_df$google_news_redirect_url, na.rm = TRUE)
      } else {
        0
      },
      if (nrow(rss_df)) n_distinct(rss_df$guid, na.rm = TRUE) else 0
    )
  )
} else {
  tibble(note = "No manifest metrics available.")
}

stage1_indicator_guide <- tribble(
  ~indicator                    , ~meaning                                                                                                              ,
  "manifest_rows"               , "One row per retrieval window tracked in the resumable manifest."                                                     ,
  "successful_windows"          , "Windows whose cached payload contains at least one RSS item."                                                        ,
  "empty_windows"               , "Windows fetched successfully but returning zero RSS items for the query."                                            ,
  "failed_or_throttled_windows" , "Windows left unfinished because the last observed attempt failed or was throttled."                                  ,
  "cached_payloads_present"     , "Manifest rows with a cached JSON payload on disk."                                                                   ,
  "completed_windows (log)"     , "Windows already complete at the time of a retrieval run, including reused cached windows from earlier runs."         ,
  "success_fetches (log)"       , "Newly fetched windows with one or more entries during that specific retrieval invocation."                           ,
  "empty_fetches (log)"         , "Newly fetched windows with zero entries during that invocation."                                                     ,
  "cached_reused (log)"         , "Previously completed windows skipped during the invocation because the cache and manifest state were already valid." ,
  "parsed_rows"                 , "Rows extracted from cached payloads before deduplication."                                                           ,
  "deduped_rows"                , "Rows retained after deterministic deduplication by redirect URL, GUID, and title+pubdate fallback."
)
```

::: {.callout-important}
The latest completed **network retrieval** run in the logs took **`r if (nrow(latest_retrieval)) latest_retrieval$elapsed[[1]] else "NA"`** for **`r if (nrow(latest_retrieval)) scales::comma(latest_retrieval$configured_windows[[1]]) else "NA"`** configured daily windows.  
The latest completed **parse + deduplication** pass took **`r if (nrow(latest_parse)) latest_parse$elapsed[[1]] else "NA"`**.

In practice, retrieval is the dominant cost. Parsing cached payloads is fast and stayed in the order of seconds on the latest run.
:::



### Detailed metrics and logs

```{r}
#| label: stage1-overview
# stage1_overview

stage1_merged <- stage1_indicator_guide |>
  left_join(stage1_overview, by = "indicator") |>
  relocate(value, .after = indicator)

kableExtra::kable(stage1_merged)
```

<!-- ### Meaning of the main indicators -->

```{r}
#| eval: false
stage1_indicator_guide
```

<!-- ### Manifest status in the current artifact -->

```{r}
#| label: stage1-manifest-status
#| eval: false
if (nrow(manifest_df)) {
  manifest_df %>%
    mutate(status = coalesce(status, "missing")) %>%
    count(status, sort = TRUE, name = "windows")
} else {
  tibble(note = "No manifest found.")
}
```

```{r}
#| eval: false
#| label: fig-stage1-status
#| fig-cap: "Stage 1 manifest status by window in the current manifest."
if (nrow(manifest_df)) {
  manifest_df %>%
    mutate(status = coalesce(status, "missing")) %>%
    count(status, sort = TRUE, name = "windows") %>%
    ggplot(aes(x = reorder(status, windows), y = windows, fill = status)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    labs(x = NULL, y = "Windows") +
    theme_minimal(base_size = 12)
}
```

<!-- ### Retrieval runtime by logged run -->

```{r}
#| label: stage1-retrieval-runtime
#| eval: false
if (nrow(retrieval_sessions)) {
  retrieval_sessions %>%
    transmute(
      run_started,
      run_finished,
      status,
      configured_start,
      configured_end,
      configured_windows,
      completed_windows,
      success_fetches,
      empty_fetches,
      cached_reused,
      failed_fetches,
      elapsed
    ) %>%
    arrange(desc(run_started))
} else {
  tibble(note = "No retrieval run history parsed from log.")
}
```

<!-- ### Parse runtime by logged run -->

```{r}
#| label: stage1-parse-runtime
#| eval: false
if (nrow(parse_sessions)) {
  parse_sessions %>%
    transmute(
      run_started,
      run_finished,
      status,
      cached_windows,
      parsed_rows,
      deduped_rows,
      missing_cache,
      elapsed
    ) %>%
    arrange(desc(run_started))
} else {
  tibble(note = "No parse run history parsed from log.")
}
```


## Step 2 Ensemble Filtering

::: {.callout-warning}
- Step 1 retrieved total 12,194 headlines. 
- The ensemble filtering stage was conducted on a sample of about **2406 headlines** (2018-2020)
:::

```{r}
#| label: stage2-overview
if (nrow(scored_df)) {
  tibble(
    metric = c(
      "headline_rows",
      "completed_rows",
      "incomplete_rows",
      "openai_error_rows",
      "local_error_rows"
    ),
    value = c(
      nrow(scored_df),
      nrow(completed_scored_df),
      nrow(scored_df) - nrow(completed_scored_df),
      sum(trimws(coalesce(scored_df$openai_error, "")) != "", na.rm = TRUE),
      sum(trimws(coalesce(scored_df$local_error, "")) != "", na.rm = TRUE)
    )
  ) %>%
    kableExtra::kable()
} else {
  tibble(note = "No Stage 2 scored output found.")
}
```

```{python}
#| label: stage2-provider-errors
from __future__ import annotations

from pathlib import Path
import pandas as pd


def find_repo_root(start: Path) -> Path:
    for path in [start, *start.parents]:
        if (path / "zika").exists():
            return path
    raise RuntimeError("Could not locate repository root.")


ROOT = find_repo_root(Path.cwd())
scored_path = ROOT / "zika" / "data" / "intermediate" / "zika_headlines_scored.csv"

if scored_path.exists():
    scored = pd.read_csv(scored_path, dtype=str).fillna("")
    error_frames = []
    for field in ["openai_error", "local_error"]:
        subset = (
            scored.loc[scored[field].str.strip() != "", field].value_counts().head(10)
        )
        if not subset.empty:
            error_frames.append(
                pd.DataFrame(
                    {
                        "field": field,
                        "error": subset.index,
                        "count": subset.values,
                    }
                )
            )

    if error_frames:
        pd.concat(error_frames, ignore_index=True)
    else:
        pd.DataFrame({"note": ["No provider errors recorded."]})
else:
    pd.DataFrame({"note": ["Scored CSV not found."]})
```

::: {.callout-caution}
While local model was able to process all 2406 headlines without error, the OpenAI batch encountered errors on 211 rows.

The error was: {"message": "Could not finish the message because max_tokens or model output limit was reached. Please try again with higher max_tokens.", "type": "invalid_request_error", "param": null, "code": null}. **Current max_tokens: 800**
:::

```{r}
#| label: stage2-agreement-helpers
stage2_label_order <- c("relevant", "irrelevant", "unsure")

stage2_paired_df <- if (nrow(completed_scored_df)) {
  completed_scored_df %>%
    filter(
      openai_label %in% stage2_label_order,
      local_label %in% stage2_label_order
    )
} else {
  tibble()
}

stage2_agreement_overview <- if (nrow(stage2_paired_df)) {
  agreement_rows <- sum(
    stage2_paired_df$openai_label == stage2_paired_df$local_label
  )
  disagreement_rows <- nrow(stage2_paired_df) - agreement_rows

  tibble(
    metric = c(
      "paired_rows",
      "agreement_rows",
      "disagreement_rows",
      "agreement_rate"
    ),
    value = c(
      nrow(stage2_paired_df),
      agreement_rows,
      disagreement_rows,
      scales::percent(agreement_rows / nrow(stage2_paired_df), accuracy = 0.1)
    )
  )
} else {
  tibble(note = "No paired OpenAI/local labels are currently available.")
}

stage2_common_share <- if (nrow(stage2_paired_df)) {
  tibble(label = stage2_label_order) %>%
    mutate(
      agreement_rows_same_label = purrr::map_int(
        label,
        ~ sum(
          stage2_paired_df$openai_label == .x &
            stage2_paired_df$local_label == .x
        )
      ),
      share_of_all_paired_rows = scales::percent(
        agreement_rows_same_label / nrow(stage2_paired_df),
        accuracy = 0.1
      )
    )
} else {
  tibble(note = "No shared-label summary available.")
}

stage2_pair_counts_long <- if (nrow(stage2_paired_df)) {
  tidyr::expand_grid(
    openai_label = stage2_label_order,
    local_label = stage2_label_order
  ) %>%
    left_join(
      stage2_paired_df %>%
        count(openai_label, local_label, name = "rows"),
      by = c("openai_label", "local_label")
    ) %>%
    mutate(
      rows = coalesce(rows, 0L),
      share_of_paired_rows = scales::percent(
        rows / nrow(stage2_paired_df),
        accuracy = 0.1
      )
    )
} else {
  tibble()
}

stage2_pair_counts_wide <- if (nrow(stage2_pair_counts_long)) {
  stage2_pair_counts_long %>%
    select(openai_label, local_label, rows) %>%
    tidyr::pivot_wider(
      names_from = local_label,
      values_from = rows
    )
} else {
  tibble(note = "No OpenAI/local cross-tab available.")
}

stage2_pair_row_share <- if (nrow(stage2_pair_counts_long)) {
  stage2_pair_counts_long %>%
    group_by(openai_label) %>%
    mutate(
      share_within_openai_label = scales::percent(
        rows / sum(rows),
        accuracy = 0.1
      )
    ) %>%
    ungroup() %>%
    select(openai_label, local_label, share_within_openai_label) %>%
    tidyr::pivot_wider(
      names_from = local_label,
      values_from = share_within_openai_label
    )
} else {
  tibble(note = "No within-OpenAI share table available.")
}
```

```{r}
#| label: stage2-model-labels
if (nrow(completed_scored_df)) {
  bind_rows(
    completed_scored_df %>%
      count(model = "OpenAI", label = openai_label, name = "rows"),
    completed_scored_df %>%
      count(model = "Local", label = local_label, name = "rows")
  ) %>%
    group_by(model) %>%
    mutate(
      share_within_model = scales::percent(rows / sum(rows), accuracy = 0.1)
    ) %>%
    ungroup() %>%
    kableExtra::kable()
} else {
  tibble(
    note = "Model label distributions are unavailable until Stage 2 completes cleanly."
  )
}
```


```{r}
#| label: stage2-agreement-overview
stage2_agreement_overview %>%
  kableExtra::kable()
```


Among the sample of **`r if (nrow(stage2_paired_df)) scales::comma(nrow(stage2_paired_df)) else "0"`** rows, there are **`r if (nrow(stage2_agreement_overview)) stage2_agreement_overview$value[stage2_agreement_overview$metric == "agreement_rows"] else 0`** rows where both models agree on the same label:

- both models called **irrelevant** on **`r if (nrow(stage2_common_share)) stage2_common_share$agreement_rows_same_label[stage2_common_share$label == "irrelevant"] else 0`** rows (**`r if (nrow(stage2_common_share)) stage2_common_share$share_of_all_paired_rows[stage2_common_share$label == "irrelevant"] else "NA"`** rows can be completely drop)
- both models called **relevant** on **`r if (nrow(stage2_common_share)) stage2_common_share$agreement_rows_same_label[stage2_common_share$label == "relevant"] else 0`** rows (**`r if (nrow(stage2_common_share)) stage2_common_share$share_of_all_paired_rows[stage2_common_share$label == "relevant"] else "NA"`** of the total sample)
- both models called **unsure** on **`r if (nrow(stage2_common_share)) stage2_common_share$agreement_rows_same_label[stage2_common_share$label == "unsure"] else 0`** rows (**`r if (nrow(stage2_common_share)) stage2_common_share$share_of_all_paired_rows[stage2_common_share$label == "unsure"] else "NA"`** of all paired rows)

**Decision rule**

- `drop` if `OpenAI = irrelevant `and `local = irrelevant` **(n=1022/1022)**.
- `keep` if `OpenAI = relevant` and `local = relevant` and rationale_similarity >= 0.75. `rationale_similarity` is only computed when both models said relevant. It uses SBERT cosine similarity on the two rationales. Otherwise it is left as not applicable **(n=614/686)**.
- `review` for everything else. That means review covers: 
  - any disagreement between models, **(n=383)**
  - any case where either model says unsure **(n=104)**
  - both say relevant but the rationale similarity is below threshold **(n=72/686, with most rationale similarities above 0.7 and just below 0.75)**.
  - both say relevant but similarity could not be computed **(n=0)**
```{r}
#| label: stage2-completed-rows
if (nrow(completed_scored_df)) {
  completed_scored_df %>%
    count(final_action, sort = TRUE, name = "rows") %>%
    kableExtra::kable()
} else {
  tibble(note = "No completed ensemble rows are currently available.")
}
```

::: {.callout-note}
Run ensemble filtering left 559 rows, about 1/4 headlines to be reviewed by human. It is important to understand the differences in how the two models approach label assignment.
:::

### OpenAI-by-local label cross-tabulation


```{r}
#| label: stage2-cross-tab-counts
stage2_pair_counts_wide %>%
  kableExtra::kable() %>%
  kableExtra::add_header_above(c(" " = 1, "deepseek-r1:14b" = 3))
```

```{r}
#| label: stage2-cross-tab-shares
stage2_pair_row_share |>
  mutate(
    relevant = ifelse(
      openai_label == "relevant",
      kableExtra::cell_spec(relevant, format = "html", background = "#d4edda"),
      as.character(relevant)
    ),
    irrelevant = ifelse(
      openai_label == "irrelevant",
      kableExtra::cell_spec(
        irrelevant,
        format = "html",
        background = "#fff3cd"
      ),
      as.character(irrelevant)
    ),
    unsure = ifelse(
      openai_label == "unsure",
      kableExtra::cell_spec(unsure, format = "html", background = "#cce5ff"),
      as.character(unsure)
    )
  ) |>
  kableExtra::kbl(
    format = "html",
    escape = FALSE,
    caption = "Within each OpenAI label, how the local model redistributed the same rows. Each row sums to 100%."
  ) |>
  kableExtra::add_header_above(c(" " = 1, "deepseek-r1:14b" = 3)) |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )
```


It is clearly seen that when OpenAI says `irrelevant`, the local model is often reluctant (just small percent-4% think that it is relevant). Instead, when OpenAI says `unsure`, the local model is only agree on 48.8% of the time, while the big difference is it thinks 32.9% of the case is `relevant`. 

Investigating further showed that OpenAI tends to require a more explicit Zika signal in the headline, whereas the local model more often treats broader vector, surveillance, or outbreak context as relevant. Some examples where OpenAI==`irrelevant` but local==`relevant`|`unsure`:

- [New Mosquito Virus Identified in Humans for the First Time Ever Dr. Axe](https://draxe.com/health/keystone-virus/)
- [Enabling Research in Low-Income Settings During Disease Outbreaks: Implementing, Learning and Preparing University of Oxford](https://news.google.com/rss/articles/CBMivgFBVV95cUxQTzNtN3F3d180ZldMWk1Ta29LRnZaNkhvb3owNHd5S3RFbzlpbmVsaE1aMlZIc29FSV81RnhuVW14S0lTN3ZZbzRBY0hRTGprRzZDZjdqZ2NpSWpqeEg4R1VRQjFaQUt2RzFkcUk4MHdnSm8xaVFFdTJUblN1RXl0LUtDZWdQLW9nYWFaZWZEZHoxLXRudjlyWlphbm52NEpveURPUVM4OTFWY3QxRVM5MmhVMEZDa1Z6Uk9KYWRR?oc=5)
- [Art and science and serendipity DePauw University](https://www.depauw.edu/stories/details/art-and-science-and-serendipity/)
- [""Helpless Means You Can't Do Anything. There's a Lot We Can Do."" Spiegel](https://www.spiegel.de/international/world/anthony-fauci-helpless-means-you-can-t-do-anyting-there-s-a-lot-we-can-do-a-211cdaf5-d0b1-41f4-b62c-25b76f4d8828)
<!-- 
The asymmetry is easiest to see in the disagreement cells:

- OpenAI `irrelevant` / local `relevant`: **`r if (nrow(stage2_pair_counts_long)) stage2_pair_counts_long$rows[stage2_pair_counts_long$openai_label == "irrelevant" & stage2_pair_counts_long$local_label == "relevant"] else 0`** rows
- OpenAI `unsure` / local `relevant`: **`r if (nrow(stage2_pair_counts_long)) stage2_pair_counts_long$rows[stage2_pair_counts_long$openai_label == "unsure" & stage2_pair_counts_long$local_label == "relevant"] else 0`** rows
- OpenAI `relevant` / local `irrelevant`: **`r if (nrow(stage2_pair_counts_long)) stage2_pair_counts_long$rows[stage2_pair_counts_long$openai_label == "relevant" & stage2_pair_counts_long$local_label == "irrelevant"] else 0`** rows -->

### Step 2 Runtime

```{r}
#| label: stage2-runtime-helpers
stage2_log_path <- file.path(ZIKA, "logs", "02_filter_headlines_ensemble.log")

parse_stage2_log <- function(path) {
  if (!file.exists(path)) {
    return(tibble())
  }

  readLines(path, warn = FALSE) %>%
    tibble(raw = .) %>%
    mutate(line_id = row_number()) %>%
    filter(str_detect(raw, "^\\d{4}-\\d{2}-\\d{2}")) %>%
    mutate(
      timestamp_txt = str_extract(
        raw,
        "^\\d{4}-\\d{2}-\\d{2} \\d{2}:\\d{2}:\\d{2}(?:[.,]\\d+)?"
      ),
      timestamp_txt = str_replace(timestamp_txt, ",", "."),
      timestamp = as.POSIXct(
        timestamp_txt,
        format = "%Y-%m-%d %H:%M:%OS",
        tz = "Europe/Stockholm"
      ),
      message = str_trim(
        str_replace(
          raw,
          "^\\d{4}-\\d{2}-\\d{2}[^|]*\\|\\s*[A-Z]+\\s*\\|\\s*",
          ""
        )
      )
    )
}

stage2_log_df <- parse_stage2_log(stage2_log_path)

openai_batch_runs <- if (nrow(stage2_log_df)) {
  openai_submissions <- stage2_log_df %>%
    filter(str_detect(message, "^Submitted OpenAI batch ")) %>%
    transmute(
      batch_id = str_match(message, "^Submitted OpenAI batch ([A-Za-z0-9_]+)")[,
        2
      ],
      submitted_at = timestamp,
      requested_rows = as.integer(str_match(message, "for (\\d+) headlines")[,
        2
      ])
    )

  openai_hydrations <- stage2_log_df %>%
    filter(str_detect(message, "^Hydrated OpenAI batch ")) %>%
    transmute(
      batch_id = str_match(message, "^Hydrated OpenAI batch ([A-Za-z0-9_]+)")[,
        2
      ],
      hydrated_at = timestamp,
      applied_rows = as.integer(str_match(message, "applied=(\\d+)")[, 2]),
      error_rows = as.integer(str_match(message, "errors=(\\d+)")[, 2])
    )

  openai_submissions %>%
    left_join(openai_hydrations, by = "batch_id") %>%
    mutate(
      turnaround_seconds = as.numeric(
        difftime(hydrated_at, submitted_at, units = "secs")
      ),
      turnaround_elapsed = vapply(
        turnaround_seconds,
        format_elapsed,
        character(1)
      )
    )
} else {
  tibble()
}

local_session_starts <- if (nrow(stage2_log_df)) {
  stage2_log_df %>%
    filter(str_detect(
      message,
      "^Stage 2 starting with \\d+ headlines; openai_pending="
    )) %>%
    transmute(
      line_id,
      run_started = timestamp,
      total_rows = as.integer(str_match(
        message,
        "^Stage 2 starting with (\\d+) headlines"
      )[, 2]),
      openai_pending = as.integer(str_match(message, "openai_pending=(\\d+)")[,
        2
      ]),
      local_pending = as.integer(str_match(message, "local_pending=(\\d+)")[,
        2
      ]),
      ensemble_incomplete = as.integer(str_match(
        message,
        "ensemble_incomplete=(\\d+)"
      )[, 2])
    )
} else {
  tibble()
}

local_scored_events <- if (nrow(stage2_log_df)) {
  stage2_log_df %>%
    filter(str_detect(message, "^Local scored ")) %>%
    transmute(
      line_id,
      scored_at = timestamp,
      progress_done = as.integer(str_match(
        message,
        "Local scored (\\d+)/(\\d+)"
      )[, 2]),
      progress_total = as.integer(str_match(
        message,
        "Local scored (\\d+)/(\\d+)"
      )[, 3])
    )
} else {
  tibble()
}

local_runtime_sessions <- if (nrow(local_session_starts)) {
  sessions <- vector("list", nrow(local_session_starts))

  for (i in seq_len(nrow(local_session_starts))) {
    next_start_line <- if (i < nrow(local_session_starts)) {
      local_session_starts$line_id[i + 1]
    } else {
      Inf
    }

    hit <- local_scored_events %>%
      filter(
        line_id > local_session_starts$line_id[i],
        line_id < next_start_line
      )

    if (!nrow(hit)) {
      sessions[[i]] <- NULL
      next
    }

    sessions[[i]] <- tibble(
      run_started = local_session_starts$run_started[i],
      local_pending = local_session_starts$local_pending[i],
      first_local_scored_at = min(hit$scored_at),
      last_local_scored_at = max(hit$scored_at),
      logged_scoring_events = nrow(hit),
      first_progress = min(hit$progress_done),
      last_progress = max(hit$progress_done),
      scoring_seconds = as.numeric(
        difftime(max(hit$scored_at), min(hit$scored_at), units = "secs")
      ),
      scoring_elapsed = format_elapsed(
        as.numeric(difftime(
          max(hit$scored_at),
          min(hit$scored_at),
          units = "secs"
        ))
      )
    )
  }

  bind_rows(sessions)
} else {
  tibble()
}

latest_successful_openai_batch <- openai_batch_runs %>%
  filter(!is.na(hydrated_at)) %>%
  arrange(desc(submitted_at)) %>%
  slice_head(n = 1)

local_runtime_summary <- if (nrow(local_runtime_sessions)) {
  active_seconds <- sum(local_runtime_sessions$scoring_seconds, na.rm = TRUE)
  first_local_at <- min(
    local_runtime_sessions$first_local_scored_at,
    na.rm = TRUE
  )
  last_local_at <- max(
    local_runtime_sessions$last_local_scored_at,
    na.rm = TRUE
  )
  wall_clock_seconds <- as.numeric(
    difftime(last_local_at, first_local_at, units = "secs")
  )
  final_local_rows <- sum(
    !is.na(scored_df$local_label) & trimws(scored_df$local_label) != "",
    na.rm = TRUE
  )

  tibble(
    metric = c(
      "local_scoring_sessions",
      "final_local_rows",
      "cumulative_active_scoring_time",
      "wall_clock_span_across_resumed_runs",
      "approx_rows_per_hour",
      "approx_seconds_per_row"
    ),
    value = c(
      nrow(local_runtime_sessions),
      final_local_rows,
      format_elapsed(active_seconds),
      format_elapsed(wall_clock_seconds),
      if (active_seconds > 0) {
        round(final_local_rows / active_seconds * 3600, 1)
      } else {
        NA_real_
      },
      if (active_seconds > 0 && final_local_rows > 0) {
        round(active_seconds / final_local_rows, 1)
      } else {
        NA_real_
      }
    )
  )
} else {
  tibble(note = "No local-scoring runtime summary available.")
}
```

```{r}
#| label: stage2-runtime-summary
if (
  nrow(latest_successful_openai_batch) &&
    all(c("metric", "value") %in% names(local_runtime_summary))
) {
  bind_rows(
    tibble(
      component = "OpenAI batch (latest successful run)",
      metric = c(
        "submitted_rows",
        "hydrated_rows",
        "error_rows",
        "turnaround_time"
      ),
      value = c(
        latest_successful_openai_batch$requested_rows,
        latest_successful_openai_batch$applied_rows,
        latest_successful_openai_batch$error_rows,
        latest_successful_openai_batch$turnaround_elapsed
      )
    ),
    local_runtime_summary %>%
      transmute(component = "Local model scoring", metric, value)
  ) %>%
    filter(metric != "wall_clock_span_across_resumed_runs") %>%
    kableExtra::kable()
} else {
  local_runtime_summary %>%
    kableExtra::kable()
}
```

- **OpenAI batch turnaround** measures elapsed time from batch submission to successful hydration of results back into the local pipeline outputs.
- **Local model runtime** measures active scoring time observed in the log across resumable Stage 2 sessions.

For the current run history, the latest successful OpenAI batch turnaround was **`r if (nrow(latest_successful_openai_batch)) latest_successful_openai_batch$turnaround_elapsed[[1]] else "NA"`**, while the local model required **`r if (nrow(local_runtime_summary)) local_runtime_summary$value[local_runtime_summary$metric == "cumulative_active_scoring_time"] else "NA"`** of active scoring time across **`r if (nrow(local_runtime_summary)) local_runtime_summary$value[local_runtime_summary$metric == "local_scoring_sessions"] else "NA"`** resumed sessions. The time needed for openAI batch is not correct and usually longer (batching often requires 24h to return outputs).


## Validation Assets

This section uses Kate's file at `zika/data/validation/zika_headline_validation_sample_full.csv` as an external reference set. The comparison is done on the intersection between the two datasets rather than on all rows. The comparison workflow is:

1. Parse and normalize Kate's file.
2. Keep only Kate rows with parseable publication dates between 2018 and 2020.
3. Match Kate rows to our experimental output using normalized Google News redirect URLs.
4. Compare Kate `RelevantYN` against our `final_action`.

```{r}
#| label: stage2-validation-helpers
clean_names_simple <- function(x) {
  x %>%
    stringr::str_replace_all("[^A-Za-z0-9]+", "_") %>%
    stringr::str_replace_all("_+", "_") %>%
    stringr::str_replace_all("^_|_$", "") %>%
    stringr::str_to_lower()
}

clean_names_unique <- function(x) {
  make.unique(clean_names_simple(x), sep = "_")
}

normalize_google_url <- function(x) {
  x <- x %>%
    coalesce("") %>%
    as.character() %>%
    stringr::str_replace_all("<[^>]+>", " ") %>%
    stringr::str_replace_all("&nbsp;", " ") %>%
    stringr::str_replace_all("&amp;", "&") %>%
    stringr::str_squish()

  extracted <- stringr::str_extract(x, "https?://[^\\s\"'>]+")
  x <- ifelse(is.na(extracted) | extracted == "", x, extracted)

  x %>%
    stringr::str_replace("[\"'>)]+$", "") %>%
    stringr::str_replace_all("(\\?|&)(hl|gl|ceid)=[^&]+", "") %>%
    stringr::str_replace_all("\\?&", "?") %>%
    stringr::str_replace_all("&&+", "&") %>%
    stringr::str_replace("[?&]+$", "")
}

binary_eval_metrics <- function(df, truth_col, pred_col, positive = "Yes") {
  if (!nrow(df)) {
    return(
      tibble(
        metric = c(
          "strict_evaluable_rows",
          "true_positive",
          "true_negative",
          "false_positive",
          "false_negative",
          "accuracy",
          "precision",
          "recall",
          "specificity",
          "f1"
        ),
        value = NA
      )
    )
  }

  truth <- df[[truth_col]]
  pred <- df[[pred_col]]

  tp <- sum(truth == positive & pred == positive, na.rm = TRUE)
  tn <- sum(truth != positive & pred != positive, na.rm = TRUE)
  fp <- sum(truth != positive & pred == positive, na.rm = TRUE)
  fn <- sum(truth == positive & pred != positive, na.rm = TRUE)
  n <- nrow(df)

  precision <- if ((tp + fp) > 0) tp / (tp + fp) else NA_real_
  recall <- if ((tp + fn) > 0) tp / (tp + fn) else NA_real_
  specificity <- if ((tn + fp) > 0) tn / (tn + fp) else NA_real_
  accuracy <- if (n > 0) (tp + tn) / n else NA_real_
  f1 <- if (!is.na(precision) && !is.na(recall) && (precision + recall) > 0) {
    2 * precision * recall / (precision + recall)
  } else {
    NA_real_
  }

  tibble(
    metric = c(
      "strict_evaluable_rows",
      "true_positive",
      "true_negative",
      "false_positive",
      "false_negative",
      "accuracy",
      "precision",
      "recall",
      "specificity",
      "f1"
    ),
    value = c(
      n,
      tp,
      tn,
      fp,
      fn,
      round(accuracy, 3),
      round(precision, 3),
      round(recall, 3),
      round(specificity, 3),
      round(f1, 3)
    )
  )
}

kate_validation_path <- file.path(
  ZIKA,
  "data",
  "validation",
  "zika_headline_validation_sample_full.csv"
)

kate_validation_raw <- safe_read_csv(kate_validation_path)

kate_validation_df <- kate_validation_raw %>%
  rename_with(clean_names_unique) %>%
  mutate(
    publication_date_std = lubridate::parse_date_time(
      publication_date,
      orders = c("ymd HMSz", "ymd HMS", "ymd HM", "ymd"),
      quiet = TRUE
    ),
    kate_google_url_norm = normalize_google_url(url),
    relevant_yn = stringr::str_to_title(stringr::str_squish(relevantyn))
  ) %>%
  filter(
    !is.na(publication_date_std),
    publication_date_std >=
      lubridate::ymd_hms("2018-01-01 00:00:00", tz = "UTC"),
    publication_date_std <
      lubridate::ymd_hms("2021-01-01 00:00:00", tz = "UTC"),
    kate_google_url_norm != ""
  ) %>%
  distinct(kate_google_url_norm, .keep_all = TRUE)

stage2_eval_df <- completed_scored_df %>%
  mutate(
    google_url_norm = normalize_google_url(google_news_redirect_url),
    final_action_binary = case_when(
      final_action == "keep" ~ "Yes",
      final_action == "drop" ~ "No",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(google_url_norm != "")

kate_overlap_df <- stage2_eval_df %>%
  inner_join(
    kate_validation_df %>%
      select(
        kate_google_url_norm,
        item_title,
        publication_date,
        relevant_yn,
        url,
        real_url,
        domain
      ),
    by = c("google_url_norm" = "kate_google_url_norm")
  ) %>%
  distinct(record_id, .keep_all = TRUE)

kate_strict_eval_df <- kate_overlap_df %>%
  filter(
    relevant_yn %in% c("Yes", "No"),
    final_action_binary %in% c("Yes", "No")
  )
```

```{r}
#| label: stage2-validation-overview
if (file.exists(kate_validation_path)) {
  tibble(
    metric = c(
      "kate_rows_total",
      "kate_rows_2018_2020_with_parseable_date",
      "stage2_completed_rows",
      "matched_common_rows",
      "matched_rows_in_review",
      "strict_evaluable_rows"
    ),
    value = c(
      nrow(kate_validation_raw),
      nrow(kate_validation_df),
      nrow(completed_scored_df),
      nrow(kate_overlap_df),
      sum(kate_overlap_df$final_action == "review", na.rm = TRUE),
      nrow(kate_strict_eval_df)
    )
  ) %>%
    kableExtra::kable()
} else {
  tibble(note = "Kate validation file not found.") %>%
    kableExtra::kable()
}
```

For the current data:

- Kate's file imports as **`r nrow(kate_validation_raw)`** rows
- only **`r nrow(kate_validation_df)`** rows have parseable publication dates between 2018 and 2020
- the normalized URL join yields **`r nrow(kate_overlap_df)`** common rows
- of those, **`r sum(kate_overlap_df$final_action == "review", na.rm = TRUE)`** marked as to be `reviewed`
- the cross evaluation is therefore uses **`r nrow(kate_strict_eval_df)`** rows
```{r}
#| label: stage2-validation-crosstab
if (nrow(kate_overlap_df)) {
  kate_overlap_df %>%
    count(relevant_yn, final_action, name = "rows") %>%
    tidyr::pivot_wider(
      names_from = final_action,
      values_from = rows,
      values_fill = 0
    ) %>%
    arrange(relevant_yn) %>%
    rename(Kate_RelevantYN = relevant_yn) %>%
    kableExtra::kable(
      # caption = "Kate RelevantYN by Stage 2 final_action on matched rows."
    ) %>%
    # add another row on top to indicate stage 2 action distribution
    kableExtra::add_header_above(
      c(" " = 1, "Ensemble filtering results" = 3)
    )
} else {
  tibble(
    note = "No matched rows between Kate's file and the Stage 2 output."
  ) %>%
    kableExtra::kable()
}
```



```{r}
#| eval: false
#| label: stage2-validation-metrics

if (nrow(kate_strict_eval_df)) {
  binary_eval_metrics(
    kate_strict_eval_df,
    truth_col = "relevant_yn",
    pred_col = "final_action_binary",
    positive = "Yes"
  ) %>%
    kableExtra::kable(
      caption = "Strict binary metrics after mapping keep = Yes and drop = No. Review rows are excluded."
    )
} else {
  tibble(note = "No strict binary evaluation rows are available.") %>%
    kableExtra::kable()
}
```

On that strict matched subset, the current ensemble action achieves:

- **accuracy = `r if (nrow(kate_strict_eval_df)) round(((sum(kate_strict_eval_df$relevant_yn == "Yes" & kate_strict_eval_df$final_action_binary == "Yes") + sum(kate_strict_eval_df$relevant_yn == "No" & kate_strict_eval_df$final_action_binary == "No")) / nrow(kate_strict_eval_df)), 3) else NA`**
- **precision = `r if (nrow(kate_strict_eval_df)) round(sum(kate_strict_eval_df$relevant_yn == "Yes" & kate_strict_eval_df$final_action_binary == "Yes") / sum(kate_strict_eval_df$final_action_binary == "Yes"), 3) else NA`**
- **recall = `r if (nrow(kate_strict_eval_df)) round(sum(kate_strict_eval_df$relevant_yn == "Yes" & kate_strict_eval_df$final_action_binary == "Yes") / sum(kate_strict_eval_df$relevant_yn == "Yes"), 3) else NA`**
- **F1 = `r if (nrow(kate_strict_eval_df)) { prec <- sum(kate_strict_eval_df$relevant_yn=="Yes" & kate_strict_eval_df$final_action_binary=="Yes") / sum(kate_strict_eval_df$final_action_binary=="Yes"); rec <- sum(kate_strict_eval_df$relevant_yn=="Yes" & kate_strict_eval_df$final_action_binary=="Yes") / sum(kate_strict_eval_df$relevant_yn=="Yes"); round(2 * prec * rec / (prec + rec), 3) } else NA`**


## Step 3 Full Text Retrieval

```{r}
#| label: stage3-runtime-helpers
parse_timestamped_log <- function(path) {
  if (!file.exists(path)) {
    return(tibble(timestamp = as.POSIXct(character()), message = character()))
  }

  readr::read_lines(path, lazy = FALSE) %>%
    tibble(line = .) %>%
    tidyr::extract(
      col = line,
      into = c("timestamp_raw", "level", "message"),
      regex = "^(\\d{4}-\\d{2}-\\d{2} \\d{2}:\\d{2}:\\d{2},\\d{3}) \\| ([A-Z]+) \\| (.*)$",
      remove = FALSE
    ) %>%
    mutate(
      timestamp = lubridate::ymd_hms(
        stringr::str_replace(timestamp_raw, ",", "."),
        tz = "Europe/Stockholm",
        quiet = TRUE
      )
    ) %>%
    filter(!is.na(timestamp))
}

safe_first_int <- function(x) {
  x <- x[!is.na(x)]
  if (!length(x)) {
    return(NA_integer_)
  }
  as.integer(x[[1]])
}

format_elapsed <- function(seconds) {
  vapply(
    seconds,
    FUN.VALUE = character(1),
    function(sec) {
      if (is.na(sec)) {
        return(NA_character_)
      }

      minutes <- floor(sec / 60)
      secs <- round(sec %% 60)
      hours <- floor(minutes / 60)
      mins <- minutes %% 60

      if (hours > 0) {
        sprintf("%dh %02dm %02ds", hours, mins, secs)
      } else if (minutes > 0) {
        sprintf("%dm %02ds", minutes, secs)
      } else {
        sprintf("%ds", secs)
      }
    }
  )
}

stage3_log_path <- file.path(ZIKA, "logs", "03_retrieve_fulltext.log")
stage3_log_df <- parse_timestamped_log(stage3_log_path)

stage3_start_df <- stage3_log_df %>%
  filter(stringr::str_detect(message, "^Stage 3 starting with ")) %>%
  tidyr::extract(
    col = message,
    into = c(
      "kept_urls_at_start",
      "pending_urls",
      "cached_success_at_start",
      "cached_failed_at_start",
      "stale_success_refetch"
    ),
    regex = paste0(
      "^Stage 3 starting with (\\d+) kept URLs",
      "(?: \\((\\d+) pending, (\\d+) cached success, (\\d+) cached failed, ",
      "(\\d+) stale success rows queued for refetch\\))?$"
    ),
    remove = FALSE
  ) %>%
  mutate(
    across(
      c(
        kept_urls_at_start,
        pending_urls,
        cached_success_at_start,
        cached_failed_at_start,
        stale_success_refetch
      ),
      ~ readr::parse_integer(.x, na = c("", "NA"))
    )
  ) %>%
  arrange(timestamp) %>%
  mutate(run_id = row_number()) %>%
  rename(run_started = timestamp)

stage3_end_df <- stage3_log_df %>%
  mutate(
    run_id = cumsum(stringr::str_detect(message, "^Stage 3 starting with "))
  ) %>%
  filter(
    stringr::str_detect(message, "^Full-text retrieval complete: ") |
      stringr::str_detect(
        message,
        "^No kept URLs found\\. Wrote empty outputs\\.$"
      )
  ) %>%
  mutate(
    new_success = readr::parse_integer(
      stringr::str_match(message, "new_success=(\\d+)")[, 2],
      na = c("", "NA")
    ),
    new_failed = readr::parse_integer(
      stringr::str_match(message, "new_failed=(\\d+)")[, 2],
      na = c("", "NA")
    ),
    total_success = readr::parse_integer(
      stringr::str_match(message, "total_success=(\\d+)")[, 2],
      na = c("", "NA")
    ),
    total_failed = readr::parse_integer(
      stringr::str_match(message, "total_failed=(\\d+)")[, 2],
      na = c("", "NA")
    ),
    legacy_success = readr::parse_integer(
      stringr::str_match(message, "success=(\\d+)")[, 2],
      na = c("", "NA")
    ),
    legacy_failed = readr::parse_integer(
      stringr::str_match(message, "failed=(\\d+)")[, 2],
      na = c("", "NA")
    ),
    no_kept_urls = stringr::str_detect(
      message,
      "^No kept URLs found\\. Wrote empty outputs\\.$"
    )
  ) %>%
  group_by(run_id) %>%
  summarise(
    run_finished = dplyr::first(timestamp),
    new_success = safe_first_int(new_success),
    new_failed = safe_first_int(new_failed),
    total_success = safe_first_int(total_success),
    total_failed = safe_first_int(total_failed),
    legacy_success = safe_first_int(legacy_success),
    legacy_failed = safe_first_int(legacy_failed),
    no_kept_urls = any(no_kept_urls),
    .groups = "drop"
  )

stage3_sessions <- stage3_start_df %>%
  left_join(stage3_end_df, by = "run_id") %>%
  mutate(
    pending_urls = dplyr::coalesce(pending_urls, kept_urls_at_start),
    processed_urls = dplyr::coalesce(new_success, legacy_success, 0L) +
      dplyr::coalesce(new_failed, legacy_failed, 0L),
    elapsed_seconds = as.numeric(difftime(
      run_finished,
      run_started,
      units = "secs"
    )),
    elapsed_time = format_elapsed(elapsed_seconds),
    approx_urls_per_minute = dplyr::if_else(
      !is.na(elapsed_seconds) & elapsed_seconds > 0,
      round(processed_urls / (elapsed_seconds / 60), 1),
      NA_real_
    )
  )

latest_stage3_fetch <- stage3_sessions %>%
  filter(processed_urls > 0, !is.na(run_finished)) %>%
  arrange(desc(run_started)) %>%
  slice_head(n = 1)
```

```{r}
#| label: stage3-runtime-summary
if (nrow(latest_stage3_fetch)) {
  tibble(
    metric = c(
      "kept_urls_at_start",
      "pending_urls",
      "new_success",
      "new_failed",
      "elapsed_time",
      "approx_urls_per_minute"
    ),
    value = c(
      latest_stage3_fetch$kept_urls_at_start[[1]],
      latest_stage3_fetch$pending_urls[[1]],
      dplyr::coalesce(
        latest_stage3_fetch$new_success[[1]],
        latest_stage3_fetch$legacy_success[[1]],
        0L
      ),
      dplyr::coalesce(
        latest_stage3_fetch$new_failed[[1]],
        latest_stage3_fetch$legacy_failed[[1]],
        0L
      ),
      latest_stage3_fetch$elapsed_time[[1]],
      latest_stage3_fetch$approx_urls_per_minute[[1]]
    )
  ) %>%
    kableExtra::kable(
      caption = "Latest substantive Step 3 retrieval run derived from the full-text log."
    )
} else {
  tibble(
    note = "No substantive full-text retrieval run was found in the Stage 3 log."
  ) %>%
    kableExtra::kable()
}
```

For the current artifacts on disk, the latest substantive Step 3 run started with **`r if (nrow(latest_stage3_fetch)) latest_stage3_fetch$kept_urls_at_start else NA`** kept URLs and **`r if (nrow(latest_stage3_fetch)) latest_stage3_fetch$pending_urls else NA`** pending fetches. It completed in **`r if (nrow(latest_stage3_fetch)) latest_stage3_fetch$elapsed_time else NA`**, producing **`r if (nrow(latest_stage3_fetch)) dplyr::coalesce(latest_stage3_fetch$new_success, latest_stage3_fetch$legacy_success, 0L) else NA`** new successful extractions and **`r if (nrow(latest_stage3_fetch)) dplyr::coalesce(latest_stage3_fetch$new_failed, latest_stage3_fetch$legacy_failed, 0L) else NA`** failed fetches. That corresponds to an observed throughput of about **`r if (nrow(latest_stage3_fetch)) latest_stage3_fetch$approx_urls_per_minute else NA`** processed URLs per minute for the most recent real retrieval pass.

### Why some full-text retrievals failed

Step 3 failures are expected in a web-scale article-retrieval pipeline because the script is operating against publisher websites with heterogeneous access controls and HTML structures. A failed row does **not** necessarily mean that the URL resolution step failed; it often means the target publisher blocked automated access, returned too little extractable article text, or timed out before the extractor could finish.

```{r}
#| label: stage3-failure-helpers
extract_domain <- function(url) {
  host <- stringr::str_match(
    url %||% "",
    "^(?:https?://)?([^/]+)"
  )[, 2]
  host <- stringr::str_to_lower(host)
  host <- stringr::str_replace(host, "^www\\.", "")
  dplyr::na_if(host, "")
}

stage3_failure_df <- if (nrow(fulltext_failed_df)) {
  fulltext_failed_df %>%
    mutate(
      failure_domain = extract_domain(final_url),
      failure_group = case_when(
        fetch_status == "http_403" ~ "publisher_blocked_http_403",
        fetch_status %in%
          c("too_short", "empty_text") ~ "insufficient_extractable_text",
        fetch_status == "request_failed" &
          stringr::str_detect(
            error_reason %||% "",
            "Read timed out"
          ) ~ "network_timeout",
        fetch_status == "request_failed" &
          stringr::str_detect(
            error_reason %||% "",
            "CERTIFICATE_VERIFY_FAILED|SSLError"
          ) ~ "ssl_or_certificate_error",
        fetch_status %in%
          c(
            "http_401",
            "http_429",
            "http_451",
            "http_500"
          ) ~ "other_http_error",
        TRUE ~ "other_failure"
      )
    )
} else {
  tibble()
}

stage3_failure_summary <- if (nrow(stage3_failure_df)) {
  stage3_failure_df %>%
    count(failure_group, sort = TRUE, name = "rows") %>%
    mutate(
      share_of_failed_rows = scales::percent(rows / sum(rows), accuracy = 0.1)
    )
} else {
  tibble()
}

stage3_failed_domain_summary <- if (nrow(stage3_failure_df)) {
  stage3_failure_df %>%
    count(failure_domain, sort = TRUE, name = "failed_rows") %>%
    slice_head(n = 10)
} else {
  tibble()
}
```

```{r}
#| label: stage3-failure-summary
if (nrow(stage3_failure_summary)) {
  stage3_failure_summary %>%
    kableExtra::kable(
      caption = "Step 3 full-text failure groups derived from the failed-output table."
    )
} else {
  tibble(note = "No failed full-text rows are present yet.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage3-failure-domains
if (nrow(stage3_failed_domain_summary)) {
  stage3_failed_domain_summary %>%
    kableExtra::kable(
      caption = "Top domains represented in the failed full-text rows."
    )
} else {
  tibble(note = "No failed full-text domains are present yet.") %>%
    kableExtra::kable()
}
```

For the current failed-output file:

- **`r if (nrow(stage3_failure_df)) sum(stage3_failure_df$fetch_status == "http_403", na.rm = TRUE) else NA` of `r if (nrow(stage3_failure_df)) nrow(stage3_failure_df) else NA`** failed rows are direct **HTTP 403** responses, which indicates that the target publisher accepted the request but refused to serve the article content to the scraper.
- **`r if (nrow(stage3_failure_df)) sum(stage3_failure_df$fetch_status %in% c("too_short", "empty_text"), na.rm = TRUE) else NA`** rows resolved successfully but produced either too little article text or no usable extraction output after the fallback extractors ran.
- **`r if (nrow(stage3_failure_df)) sum(stage3_failure_df$failure_group == "network_timeout", na.rm = TRUE) else NA`** rows failed because the destination site timed out before the page could be fetched, and **`r if (nrow(stage3_failure_df)) sum(stage3_failure_df$failure_group == "ssl_or_certificate_error", na.rm = TRUE) else NA`** row failed because of an SSL certificate verification problem.

The main interpretation is that Step 3 is constrained more by **publisher access control and network variability** than by the extraction code itself. In the current run, the single biggest source of failure is publisher blocking via HTTP 403, while the next tier of failures comes from pages that either timed out or yielded text that was too short to count as a usable article body.

## Step 4 BERTopic on the retrieved full-text subset

To explore whether the retained full-text corpus contains coherent thematic structure beyond the headline labels, I first ran a standard BERTopic pipeline over the successful Stage 3 subset and then added an **impact-guided** BERTopic refinement. The modeling input is intentionally narrower than the raw `zika_fulltext.csv` table: the script keeps documents with non-empty text, enforces a minimum word-count threshold, applies a lightweight English-language filter, and truncates each document to a fixed word budget before embedding.

That refinement matters for interpretation. The fully unsupervised MiniLM baseline tended to collapse the retained corpus into very broad Zika-centered clusters, which is not especially useful when the research goal is to characterize **impact pathways**. The new impact-guided run therefore suppresses generic Zika vocabulary, seeds BERTopic with impact-oriented concept bundles, and uses a KMeans partition to force a more inspectable thematic split.

Following BERTopic's modular design, the current report now treats the pipeline as six separate submodels:

1. embeddings
2. dimensionality reduction
3. clustering
4. vectorization
5. c-TF-IDF weighting
6. representation modeling

The generic `zika_bertopic_*` outputs correspond to the current **baseline** embedding backend, which remains `all-MiniLM-L6-v2`. The richer modular outputs are stored separately in `zika_bertopic_modular_analysis.*`, `zika_bertopic_modular_topics.csv`, and `zika_bertopic_stage_timings.csv`. The new impact-guided outputs are written to `zika_bertopic_impact_*`.

```{r}
#| label: stage4-bertopic-overview
if (length(bertopic_impact_payload)) {
  tibble(
    metric = c(
      "documents_input",
      "documents_modeled",
      "documents_excluded",
      "unique_topics_excluding_outlier",
      "outlier_documents",
      "embedding_seconds",
      "bertopic_fit_seconds",
      "projection_seconds",
      "runtime_human"
    ),
    value = c(
      bertopic_impact_payload$documents_input %||% NA,
      bertopic_impact_payload$documents_modeled %||% NA,
      bertopic_impact_payload$documents_excluded %||% NA,
      bertopic_impact_payload$unique_topics_excluding_outlier %||% NA,
      bertopic_impact_payload$outlier_documents %||% NA,
      bertopic_impact_payload$stage_timings$embedding_encode %||% NA,
      bertopic_impact_payload$stage_timings$bertopic_fit_transform %||% NA,
      bertopic_impact_payload$stage_timings$projection_umap_2d %||% NA,
      bertopic_impact_payload$runtime_human %||% NA
    )
  ) %>%
    kableExtra::kable(
      caption = "Impact-guided BERTopic run summary for the current full-text subset."
    )
} else {
  tibble(note = "Impact-guided BERTopic outputs are not present yet.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage4-bertopic-component-specification
tibble(
  pipeline_step = c(
    "1. Embeddings",
    "2. Dimensionality reduction",
    "3. Clustering",
    "4. Vectorizer",
    "5. c-TF-IDF",
    "6. Representation model"
  ),
  baseline_choice = c(
    "all-MiniLM-L6-v2 and BAAI/bge-base-en-v1.5",
    "UMAP (n_neighbors = 15, n_components = 5, min_dist = 0, cosine metric)",
    "HDBSCAN (min_cluster_size = 12, eom)",
    "CountVectorizer with English stop words, min_df = 2, ngrams = 1:2",
    "ClassTfidfTransformer(reduce_frequent_words = TRUE)",
    "KeyBERTInspired"
  ),
  analysis_added_in_report = c(
    "Compare MiniLM versus BGE on the same filtered corpus.",
    "Compare UMAP against PCA while holding the primary embedding fixed.",
    "Compare HDBSCAN against KMeans while reusing the same MiniLM embeddings.",
    "Compare basic unigram versus tuned bigram CountVectorizer on fixed topic assignments.",
    "Compare standard, reduce-frequent, and BM25-plus-reduce-frequent weighting.",
    "Compare c-TF-IDF only, KeyBERTInspired, and MMR representations."
  )
) %>%
  kableExtra::kable(
    caption = "BERTopic submodels used in the current analysis and the explicit ablations added for this report."
  )
```

```{r}
#| label: stage4-impact-helpers
impact_topic_label <- function(topic_name, representation) {
  purrr::map2_chr(
    topic_name,
    representation,
    function(topic_name_i, representation_i) {
      text <- paste(topic_name_i %||% "", representation_i %||% "")
      text <- tolower(text)

      dplyr::case_when(
        str_detect(text, "antiviral|antibod|vaccine|flavivirus|proteins") ~
          "Vaccines and therapeutics",
        str_detect(text, "microcephaly|birth defects|children|developmental") ~
          "Congenital outcomes and child risk",
        str_detect(text, "pregnancy|congenital|neural|stem cells|fetuses") ~
          "Pregnancy and fetal neurodevelopment",
        str_detect(
          text,
          "vector|climate change|aedes|mosquito species|albopictus"
        ) ~
          "Vector ecology and climate risk",
        str_detect(text, "caribbean|travelers|cuba|islands|public health") ~
          "Travel advisories and regional surveillance",
        str_detect(text, "symptoms|testing|rash|pregnant women|rajasthan") ~
          "Clinical symptoms and public guidance",
        str_detect(text, "assay|pcr|diagnostic|samples|blood supply") ~
          "Diagnostics and laboratory surveillance",
        str_detect(
          text,
          "semen|reproductive|testis|sexually transmitted|intravaginal"
        ) ~
          "Sexual and reproductive transmission",
        TRUE ~ str_replace_all(topic_name_i %||% "Unlabeled topic", "_", " ")
      )
    }
  )
}

impact_topic_lookup_df <- if (nrow(bertopic_impact_topic_info_df)) {
  bertopic_impact_topic_info_df %>%
    filter(Topic != -1) %>%
    transmute(
      topic_id = as.integer(Topic),
      document_count = as.integer(Count),
      topic_name = as.character(Name),
      top_terms = as.character(Representation),
      impact_label = impact_topic_label(topic_name, top_terms)
    ) %>%
    arrange(desc(document_count), topic_id)
} else {
  tibble()
}

impact_scored_join_df <- if ("record_id" %in% names(scored_df)) {
  scored_df %>%
    select(any_of(c("record_id", "rss_pubdate")))
} else {
  tibble(record_id = character(), rss_pubdate = character())
}

impact_fulltext_join_df <- if ("record_id" %in% names(fulltext_df)) {
  fulltext_df %>%
    select(any_of(c("record_id", "published_at")))
} else {
  tibble(record_id = character(), published_at = character())
}

impact_doc_enriched_df <- if (nrow(bertopic_impact_doc_df)) {
  bertopic_impact_doc_df %>%
    left_join(
      impact_topic_lookup_df,
      by = c("topic_id", "topic_name")
    ) %>%
    left_join(impact_scored_join_df, by = "record_id") %>%
    left_join(impact_fulltext_join_df, by = "record_id") %>%
    mutate(
      rss_pubdate_dt = suppressWarnings(
        lubridate::parse_date_time(
          rss_pubdate,
          orders = c("a, d b Y H:M:S z", "ymd HMS", "ymd", "Ymd HMS", "Ymd")
        )
      ),
      published_at_dt = suppressWarnings(
        lubridate::parse_date_time(
          published_at,
          orders = c("ymd HMS", "ymd HM", "ymd", "Ymd HMS", "Ymd")
        )
      ),
      event_date = coalesce(as.Date(published_at_dt), as.Date(rss_pubdate_dt)),
      event_year = lubridate::year(event_date)
    )
} else {
  tibble()
}

impact_projection_plot_df <- if (nrow(bertopic_impact_projection_df)) {
  bertopic_impact_projection_df %>%
    left_join(
      impact_topic_lookup_df %>%
        select(topic_id, topic_name, impact_label),
      by = c("topic_id", "topic_name")
    )
} else {
  tibble()
}

impact_topic_size_plot_df <- if (nrow(impact_topic_lookup_df)) {
  impact_topic_lookup_df %>%
    mutate(
      share_of_modeled_docs = document_count / sum(document_count),
      impact_label = forcats::fct_reorder(impact_label, document_count)
    )
} else {
  tibble()
}

impact_topic_timeline_df <- if (nrow(impact_doc_enriched_df)) {
  impact_doc_enriched_df %>%
    filter(!is.na(event_year)) %>%
    count(event_year, impact_label, name = "documents") %>%
    group_by(event_year) %>%
    mutate(share_within_year = documents / sum(documents)) %>%
    ungroup()
} else {
  tibble()
}

impact_representative_docs_df <- if (nrow(impact_doc_enriched_df)) {
  impact_doc_enriched_df %>%
    arrange(impact_label, desc(model_word_count), rss_title) %>%
    group_by(impact_label) %>%
    slice_head(n = 3) %>%
    ungroup() %>%
    transmute(
      impact_label,
      year = event_year,
      headline = rss_title,
      domain,
      words = model_word_count
    )
} else {
  tibble()
}

impact_runtime_df <- if (nrow(bertopic_impact_stage_timing_df)) {
  bertopic_impact_stage_timing_df %>%
    mutate(
      stage_name = recode(
        stage_name,
        embedding_encode = "Embedding",
        bertopic_fit_transform = "BERTopic fit",
        projection_umap_2d = "2D projection"
      )
    )
} else {
  tibble()
}
```

```{r}
#| label: stage4-bertopic-compare-helpers
bertopic_pairwise_df <- if (
  !is.null(bertopic_compare_payload$pairwise_metrics) &&
    length(bertopic_compare_payload$pairwise_metrics)
) {
  tibble::as_tibble(bertopic_compare_payload$pairwise_metrics)
} else {
  tibble()
}

bertopic_modular_display_df <- if (nrow(bertopic_modular_df)) {
  bertopic_modular_df %>%
    mutate(
      analysis_group = factor(
        analysis_group,
        levels = c(
          "dimensionality_reduction",
          "clustering",
          "vectorizer",
          "ctfidf",
          "representation"
        )
      ),
      largest_topic_share_pct = 100 * largest_topic_share,
      runtime_seconds = round(runtime_seconds, 3),
      normalized_topic_entropy = round(normalized_topic_entropy, 3),
      mean_topic_ctfidf_cosine = round(mean_topic_ctfidf_cosine, 3)
    )
} else {
  tibble()
}

bertopic_runtime_plot_df <- bind_rows(
  if (nrow(bertopic_compare_df)) {
    bertopic_compare_df %>%
      transmute(
        analysis_group = "embedding_model",
        analysis_variant = embedding_model,
        metric = "Embedding time (s)",
        value = embedding_seconds
      )
  },
  if (nrow(bertopic_compare_df)) {
    bertopic_compare_df %>%
      transmute(
        analysis_group = "embedding_model",
        analysis_variant = embedding_model,
        metric = "BERTopic fit time (s)",
        value = analysis_seconds
      )
  },
  if (nrow(bertopic_modular_display_df)) {
    bertopic_modular_display_df %>%
      transmute(
        analysis_group = as.character(analysis_group),
        analysis_variant,
        metric = "Total runtime (s)",
        value = runtime_seconds
      )
  }
) %>%
  mutate(
    analysis_group = factor(
      analysis_group,
      levels = c(
        "embedding_model",
        "dimensionality_reduction",
        "clustering",
        "vectorizer",
        "ctfidf",
        "representation"
      )
    ),
    analysis_variant = forcats::fct_inorder(analysis_variant)
  )

bertopic_runtime_table_df <- if (nrow(bertopic_stage_timing_df)) {
  bertopic_stage_timing_df %>%
    mutate(
      analysis_group = factor(
        analysis_group,
        levels = c(
          "dimensionality_reduction",
          "clustering",
          "vectorizer",
          "ctfidf",
          "representation"
        )
      )
    ) %>%
    arrange(analysis_group, analysis_variant, stage_name)
} else {
  tibble()
}

bertopic_topic_size_plot_df <- {
  raw_topic_size_df <- bind_rows(
    if (nrow(bertopic_minilm_topic_info_df)) {
      bertopic_minilm_topic_info_df %>%
        mutate(embedding_model = "all-MiniLM-L6-v2")
    },
    if (nrow(bertopic_bge_topic_info_df)) {
      bertopic_bge_topic_info_df %>%
        mutate(embedding_model = "BAAI/bge-base-en-v1.5")
    }
  )

  if (nrow(raw_topic_size_df)) {
    raw_topic_size_df %>%
      filter(Topic != -1) %>%
      transmute(
        embedding_model,
        topic_id = as.integer(Topic),
        document_count = as.integer(Count),
        topic_label = paste0("T", topic_id)
      ) %>%
      group_by(embedding_model) %>%
      mutate(
        share_of_model = document_count / sum(document_count),
        topic_label = forcats::fct_reorder(topic_label, document_count)
      ) %>%
      ungroup()
  } else {
    tibble()
  }
}

bertopic_compare_plot_df <- if (nrow(bertopic_compare_df)) {
  bertopic_compare_df %>%
    transmute(
      embedding_model,
      `Topics` = unique_topics_excluding_outlier,
      `Outliers` = outlier_documents,
      `Largest topic (%)` = 100 * largest_topic_share,
      `Embedding time (s)` = embedding_seconds,
      `BERTopic fit (s)` = analysis_seconds,
      `Topic entropy` = normalized_topic_entropy
    ) %>%
    tidyr::pivot_longer(
      cols = -embedding_model,
      names_to = "metric",
      values_to = "value"
    )
} else {
  tibble()
}

bertopic_assignment_heatmap_df <- if (
  nrow(bertopic_minilm_doc_df) && nrow(bertopic_bge_doc_df)
) {
  bertopic_minilm_doc_df %>%
    select(record_id, minilm_topic = topic_id) %>%
    inner_join(
      bertopic_bge_doc_df %>%
        select(record_id, bge_topic = topic_id),
      by = "record_id"
    ) %>%
    mutate(
      minilm_topic = if_else(
        minilm_topic == -1,
        "outlier",
        paste0("T", minilm_topic)
      ),
      bge_topic = if_else(bge_topic == -1, "outlier", paste0("T", bge_topic))
    ) %>%
    count(minilm_topic, bge_topic, name = "documents")
} else {
  tibble()
}
```

### Impact-guided topic structure

```{r}
#| label: stage4-impact-topic-summary
if (nrow(impact_topic_lookup_df)) {
  impact_topic_lookup_df %>%
    mutate(
      `share of modeled docs` = scales::percent(
        document_count / sum(document_count),
        accuracy = 0.1
      )
    ) %>%
    select(
      topic_id,
      impact_label,
      document_count,
      `share of modeled docs`,
      top_terms
    ) %>%
    kableExtra::kable(
      caption = "Impact-guided BERTopic topics after removing generic Zika vocabulary and seeding the model with impact-oriented concept bundles."
    )
} else {
  tibble(note = "No impact-guided BERTopic topics are available.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage4-impact-runtime
if (nrow(impact_runtime_df)) {
  impact_runtime_df %>%
    select(stage_name, runtime_seconds, runtime_human) %>%
    kableExtra::kable(
      caption = "Runtime required for the impact-guided BERTopic analysis."
    )
} else {
  tibble(note = "No impact-guided runtime table is available.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: fig-stage4-impact-topic-size
#| fig-cap: "Document counts by impact-guided BERTopic topic. This view is intended to show which impact pathways dominate the retained full-text corpus."
#| fig-width: 9.5
#| fig-height: 5.5
if (nrow(impact_topic_size_plot_df)) {
  ggplot(
    impact_topic_size_plot_df,
    aes(
      x = impact_label,
      y = document_count,
      fill = share_of_modeled_docs
    )
  ) +
    geom_col(width = 0.75) +
    geom_text(
      aes(label = document_count),
      hjust = -0.15,
      size = 3.1
    ) +
    coord_flip(clip = "off") +
    scale_fill_gradient(
      low = "#d9ed92",
      high = "#1d3557",
      labels = scales::percent_format(accuracy = 1)
    ) +
    expand_limits(y = max(impact_topic_size_plot_df$document_count) * 1.08) +
    labs(
      x = NULL,
      y = "Documents",
      fill = "Share of modeled docs"
    ) +
    theme_minimal(base_size = 11) +
    theme(
      legend.position = "bottom",
      panel.grid.major.y = element_blank()
    )
}
```

```{r}
#| label: fig-stage4-impact-projection
#| fig-cap: "Two-dimensional UMAP projection of the impact-guided BERTopic assignments. Nearby points have similar embeddings; color indicates the impact theme assigned by BERTopic."
#| fig-width: 9
#| fig-height: 6
if (nrow(impact_projection_plot_df)) {
  ggplot(
    impact_projection_plot_df,
    aes(
      x = x,
      y = y,
      color = impact_label
    )
  ) +
    geom_point(alpha = 0.72, size = 1.8) +
    labs(
      x = "UMAP 1",
      y = "UMAP 2",
      color = NULL
    ) +
    theme_minimal(base_size = 11) +
    theme(
      legend.position = "bottom",
      panel.grid = element_blank()
    )
}
```

```{r}
#| label: fig-stage4-impact-timeline
#| fig-cap: "Year-by-year topic mix in the impact-guided BERTopic run. Within each year, the stacked bars show the share of documents allocated to each impact theme."
#| fig-width: 10
#| fig-height: 5.5
if (nrow(impact_topic_timeline_df)) {
  ggplot(
    impact_topic_timeline_df,
    aes(
      x = factor(event_year),
      y = share_within_year,
      fill = impact_label
    )
  ) +
    geom_col(width = 0.8) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    labs(
      x = "Year",
      y = "Share of modeled docs within year",
      fill = NULL
    ) +
    theme_minimal(base_size = 11) +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 35, hjust = 1)
    )
}
```

```{r}
#| label: stage4-impact-representative-docs
if (nrow(impact_representative_docs_df)) {
  impact_representative_docs_df %>%
    kableExtra::kable(
      caption = "Representative headlines from the impact-guided BERTopic topics. Within each topic, examples are the longest modeled documents in the retained full-text subset."
    )
} else {
  tibble(note = "No representative impact-topic headlines are available.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage4-bertopic-runtime-table
if (nrow(bertopic_runtime_table_df)) {
  bertopic_runtime_table_df %>%
    select(
      analysis_group,
      analysis_variant,
      stage_name,
      runtime_seconds,
      runtime_human
    ) %>%
    kableExtra::kable(
      caption = "Measured runtime by BERTopic analysis. Dimensionality-reduction and clustering baselines include the MiniLM embedding pass; later lexical ablations time only `update_topics()`."
    )
} else {
  tibble(note = "No BERTopic stage timing table is available.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage4-bertopic-runtime-figure
#| fig-cap: "Runtime required for each BERTopic analysis. Embedding comparisons are split into encoding and BERTopic fit time; later module ablations reuse cached MiniLM embeddings."
#| fig-width: 10
#| fig-height: 6
if (nrow(bertopic_runtime_plot_df)) {
  ggplot(
    bertopic_runtime_plot_df,
    aes(
      x = analysis_variant,
      y = value,
      fill = metric
    )
  ) +
    geom_col(width = 0.72) +
    facet_wrap(~analysis_group, scales = "free_x") +
    scale_fill_manual(
      values = c(
        "Embedding time (s)" = "#355070",
        "BERTopic fit time (s)" = "#6d597a",
        "Total runtime (s)" = "#b56576"
      )
    ) +
    labs(
      x = NULL,
      y = "Seconds",
      fill = NULL
    ) +
    theme_minimal(base_size = 11) +
    theme(
      legend.position = "bottom",
      axis.text.x = element_text(angle = 25, hjust = 1),
      strip.text = element_text(face = "bold")
    )
}
```

```{r}
#| label: stage4-bertopic-topic-size-figure
#| fig-cap: "Topic-size structure under each embedding backend. MiniLM yields a highly concentrated solution, while BGE splits the corpus into more subtopics."
#| fig-width: 9
#| fig-height: 5
if (nrow(bertopic_topic_size_plot_df)) {
  ggplot(
    bertopic_topic_size_plot_df,
    aes(
      x = topic_label,
      y = document_count,
      fill = share_of_model
    )
  ) +
    geom_col(width = 0.75) +
    geom_text(
      aes(label = document_count),
      hjust = -0.1,
      size = 3
    ) +
    coord_flip(clip = "off") +
    facet_wrap(~embedding_model, scales = "free_y") +
    scale_fill_gradient(
      low = "#9ecae1",
      high = "#08519c",
      labels = scales::percent_format(accuracy = 1)
    ) +
    expand_limits(y = max(bertopic_topic_size_plot_df$document_count) * 1.08) +
    labs(
      x = NULL,
      y = "Documents",
      fill = "Share within model"
    ) +
    theme_minimal(base_size = 11) +
    theme(
      legend.position = "bottom",
      panel.grid.major.y = element_blank()
    )
}
```

```{r}
#| label: stage4-bertopic-model-comparison-figure
#| fig-cap: "Model-level BERTopic comparison across MiniLM and BGE on the same filtered full-text subset."
#| fig-width: 9
#| fig-height: 5.5
if (nrow(bertopic_compare_plot_df)) {
  ggplot(
    bertopic_compare_plot_df,
    aes(
      x = embedding_model,
      y = value,
      fill = embedding_model
    )
  ) +
    geom_col(width = 0.65, show.legend = FALSE) +
    geom_text(
      aes(label = round(value, 2)),
      vjust = -0.3,
      size = 3
    ) +
    facet_wrap(~metric, scales = "free_y") +
    scale_fill_manual(
      values = c(
        "all-MiniLM-L6-v2" = "#264653",
        "BAAI/bge-base-en-v1.5" = "#e76f51"
      )
    ) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(
      panel.grid.major.x = element_blank(),
      strip.text = element_text(face = "bold")
    )
}
```

```{r}
#| label: stage4-bertopic-assignment-heatmap
#| fig-cap: "Document-level reassignment from MiniLM topics to BGE topics. Darker cells indicate more documents following that path."
#| fig-width: 7.5
#| fig-height: 5.5
if (nrow(bertopic_assignment_heatmap_df)) {
  ggplot(
    bertopic_assignment_heatmap_df,
    aes(
      x = minilm_topic,
      y = bge_topic,
      fill = documents
    )
  ) +
    geom_tile(color = "white", linewidth = 0.4) +
    geom_text(aes(label = documents), size = 3) +
    scale_fill_gradient(low = "#fef6e4", high = "#c1121f") +
    labs(
      x = "MiniLM topic",
      y = "BGE topic",
      fill = "Documents"
    ) +
    theme_minimal(base_size = 11) +
    theme(
      panel.grid = element_blank(),
      axis.text.x = element_text(angle = 30, hjust = 1)
    )
}
```

```{r}
#| label: stage4-bertopic-topic-table
if (nrow(bertopic_topic_info_df)) {
  bertopic_topic_info_df %>%
    transmute(
      topic_id = Topic,
      document_count = Count,
      topic_name = Name,
      top_terms = Representation
    ) %>%
    arrange(topic_id) %>%
    kableExtra::kable(
      caption = "Primary BERTopic topic summary. Topic `-1` would indicate outliers; none were produced in the current MiniLM run."
    )
} else {
  tibble(note = "No BERTopic topic table is available.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage4-bertopic-document-shares
if (nrow(bertopic_doc_df)) {
  bertopic_doc_df %>%
    count(topic_id, topic_name, name = "documents") %>%
    mutate(
      share_of_modeled_docs = scales::percent(
        documents / sum(documents),
        accuracy = 0.1
      )
    ) %>%
    arrange(desc(documents), topic_id) %>%
    kableExtra::kable(
      caption = "Document share by BERTopic topic in the modeled subset."
    )
} else {
  tibble(note = "No BERTopic document-topic assignments are available.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage4-bertopic-model-comparison
if (nrow(bertopic_compare_df)) {
  bertopic_compare_df %>%
    mutate(
      largest_topic_share = scales::percent(
        largest_topic_share,
        accuracy = 0.1
      ),
      normalized_topic_entropy = round(normalized_topic_entropy, 3),
      mean_topic_ctfidf_cosine = round(mean_topic_ctfidf_cosine, 3),
      max_topic_ctfidf_cosine = round(max_topic_ctfidf_cosine, 3)
    ) %>%
    arrange(embedding_model) %>%
    kableExtra::kable(
      caption = "Per-model BERTopic comparison using the same filtered corpus."
    )
} else {
  tibble(note = "No BERTopic model-comparison table is available.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage4-bertopic-modular-summary
if (nrow(bertopic_modular_display_df)) {
  bertopic_modular_display_df %>%
    transmute(
      analysis_group,
      analysis_variant,
      runtime_seconds,
      topics = unique_topics_excluding_outlier,
      outliers = outlier_documents,
      `largest topic (%)` = round(largest_topic_share_pct, 1),
      `topic entropy` = normalized_topic_entropy,
      `mean c-TF-IDF cosine` = mean_topic_ctfidf_cosine,
      `unique representation terms` = unique_representation_terms
    ) %>%
    arrange(analysis_group, runtime_seconds) %>%
    kableExtra::kable(
      caption = "Summary of the modular BERTopic analyses on the primary MiniLM embedding run."
    )
} else {
  tibble(note = "No BERTopic modular analysis summary is available.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage4-bertopic-pairwise
if (nrow(bertopic_pairwise_df)) {
  bertopic_pairwise_df %>%
    kableExtra::kable(
      caption = "Pairwise agreement between embedding backends on document-level topic assignments."
    )
} else {
  tibble(note = "No BERTopic pairwise comparison metrics are available.") %>%
    kableExtra::kable()
}
```

```{r}
#| label: stage4-bertopic-topic-matches
if (nrow(bertopic_topic_match_df)) {
  bertopic_topic_match_df %>%
    arrange(source_model, desc(ctfidf_cosine_similarity), source_topic_id) %>%
    kableExtra::kable(
      caption = "Best cross-model topic matches using post-hoc c-TF-IDF cosine similarity on the same topic pseudo-documents."
    )
} else {
  tibble(note = "No BERTopic cross-model topic matches are available.") %>%
    kableExtra::kable()
}
```

For the current **impact-guided** run, BERTopic received **`r bertopic_impact_payload$documents_input %||% NA`** successful full-text rows and modeled **`r bertopic_impact_payload$documents_modeled %||% NA`** documents after filtering. The full impact analysis took **`r bertopic_impact_payload$runtime_human %||% NA`**, which splits into about **`r round(bertopic_impact_payload$stage_timings$embedding_encode %||% NA, 1)` seconds** for embedding generation, **`r round(bertopic_impact_payload$stage_timings$bertopic_fit_transform %||% NA, 1)` seconds** for BERTopic fitting, and **`r round(bertopic_impact_payload$stage_timings$projection_umap_2d %||% NA, 1)` seconds** for the 2D inspection layout.

Substantively, the impact-guided model is much closer to the research objective than the coarse unsupervised baseline. Instead of returning one dominant “general Zika” topic, it separates the retained corpus into interpretable impact pathways such as:

- congenital outcomes and child risk
- pregnancy and fetal neurodevelopment
- vector ecology and climate risk
- diagnostics and laboratory surveillance
- travel advisories and regional surveillance
- sexual and reproductive transmission
- vaccines and therapeutics

This shift comes from three deliberate modeling choices: removing generic Zika vocabulary from the vectorizer, seeding the model with impact-oriented concept bundles, and using KMeans to force a finer topic partition. On the current corpus, those choices produce **`r bertopic_impact_payload$unique_topics_excluding_outlier %||% NA`** non-outlier topics with **no outlier bucket**, and the largest topic accounts for only **`r scales::percent(bertopic_impact_payload$largest_topic_share %||% NA, accuracy = 0.1)`** of modeled documents. That is a much more useful structure for downstream impact screening than the earlier MiniLM baseline, where one topic dominated nearly the entire corpus.

The visual diagnostics help show why the new result is easier to investigate:

- the topic-size chart shows that the retained full-text subset is no longer collapsing into one overwhelmingly generic theme
- the 2D projection reveals distinct neighborhoods for reproductive transmission, diagnostics, travel/public-health alerts, and congenital outcomes
- the yearly topic-mix figure makes it easier to see when certain impact pathways become more prevalent in the corpus

The earlier embedding and modular comparisons are still useful methodologically. They explain why the pipeline was refined in this direction:

- **Embedding comparison:** `all-MiniLM-L6-v2` remains faster and more stable than `BAAI/bge-base-en-v1.5`, while BGE produces a more fragmented solution with **`r if (nrow(bertopic_compare_df)) bertopic_compare_df$outlier_documents[bertopic_compare_df$embedding_model == "BAAI/bge-base-en-v1.5"][[1]] else NA`** outliers.
- **Dimensionality reduction:** UMAP is slower than PCA, but PCA still produces many outliers and a flatter, less interpretable structure.
- **Clustering:** KMeans is the key reason the impact-guided run becomes granular enough to be useful. In the modular comparison it produces a finer partition than HDBSCAN on the same MiniLM embeddings.
- **Vectorizer, c-TF-IDF, and representation:** these layers mostly affect topic wording, but once generic Zika words are suppressed they help surface more specific impact terms such as `microcephaly`, `pregnancy`, `pcr`, `caribbean`, and `semen`.

So the practical interpretation is:

- the original unsupervised BERTopic baseline is still helpful as a methodological reference, but it is too generic for the impact-filtering objective
- the new impact-guided BERTopic run is the more useful analytical output for this project because it organizes the retained Zika corpus into concrete impact pathways rather than repeating that the documents are about Zika
- the inspection figures make it easier to decide which topics are worth keeping, merging, or refining further in later iterations
